### 2025-01-29 products/manuscript/Manuscript.qmd
Part 1 of MADA Project

---
title: "Immune-Related Adverse Events (irAEs) associated with Immune Checkpoint Inhibitor (ICI) Therapy in Bladder Cancer: An Analysis of the FDA Adverse Event Reporting System (FAERS)"
author: "Pooja Gokhale, PharmD"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/nlm-style.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

**Authors**

-   Pooja Gokhale, PharmD$^{1}$ (ORCID: 0000-0002-7505-7007)

**Author affiliations**

1.  Department of Clinical and Administrative Pharmacy, College of Pharmacy, University of Georgia, Athens, GA, USA.

$\land$ Corresponding author: poojaGokhale\@uga.edu

{{< pagebreak >}}

# Summary/Abstract

*Write a summary of your project.*
*This part has not been completed. The source code has been left as is*

{{< pagebreak >}}

# Introduction

## General Background Information

*Bladder cancer is the ninth most common type of cancer, according to a 2022 report published by the International Agency for Research on Cancer [@Bray2024]. The American Cancer Society estimates approximately 85,000 new bladder cancer cases, and approximately 17,000 bladder cancer deaths in the United States in 2025.*

*The cancer is more common in males than in females, with a male:female ratio of about 4:1[@Leslie2025-lp]. Risk factors for bladder cancer include smoking and occupational chemical exposures (in particular, aromatic amines and polycyclic aromatic hydrocarbons). Certain dietary factors, impaired microbiome, and pelvic radiotherapy may also increase the risk of bladder cancer [@Jubber2023].*

*Treatment of bladder cancer is multifaceted and depends on the type of tumor. Surgical resection, adjuvant and intravesical chemotherapy, Bacillus Calmette–Guérin (BCG) immunotherapy, and systemic chemotherapy are the different treatment options available [@Leslie2025-lp].*

*According to the National comprehensive Cancer Network (NCCN) guidelines, immunotherapy with immune checkpoint inhibitors (ICIs) is recommended at various stages of bladder cancer treatment. ICIs such as pembrolizumab, atezolizumab, nivolumab, and avelumab are FDA approved and recommended for the adjuvant and first and second-line treatment of bladder cancer [@Flaig2024].*

*ICIs are associated with various adverse effects, particularly with immune-related adverse events (irAEs) such as pnemonitis, hepatitis, colitis, hypothyroidism, dermatitis, and others [@Yin2023].This study aims to analyze real-world adverse event data from the FDA Adverse Event Reporting System associated with ICI therapy in bladder cancer.*

## Description of data and data source

*The United States Food and Drug Administration (FDA) maintains a dashboard of adverse event reports in the FDA Adverse Event Reporting System (FAERS) [@us2018fda]. FAERS is a publicly available dashboard of safety reports submitted by patients, healthcare professionals, and pharmaceutical companies.*

*Since FAERS consists of self-reported data from multiple sources, there may be duplicates, missing data as well as various errors. The findings of this study should be considered in context of this data limitation.*

## Questions/Hypotheses to be addressed

*The study aims to analyze whether irAEs associated with ICIs generate a signal of disproportionate reporting (SDR). The goal is to determine whether irAEs associated with ICIs in baldder cancer are greater than expected, and thus further research and strategies to deal with these adverse events is required.*

{{< pagebreak >}}

# Methods

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*
*This part has not been completed. The source code has been left as is*
*Details have been provided in the data acquisition and statistical analysis parts.*

## Schematic of workflow

*This part has not been completed. The source code has been left as is*

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender. We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition

*In order to obtain the FAERS data, the ICIs were used to search by product. Pembrolizumab, atezolizumab, avelumab, and nivolumab were the keywords used. Previously, durvalumab, and ipilimumab were also approved for bladder cancer treatment. However, this indication has since then been removed by the manufacturers. As such, these were not considered. Following this, the search was limited to cases of bladder cancer. A total of 1,886 reactions were obtained.*

## Data import and cleaning

*The file has been downloaded as an Excel file. The code for the analysis would involve loading the file, removing duplicates, assessing missing data, and deciding whether to remove it.* *This will be followed by filtering the data to retain only irAEs. While this step could be performed within the FAERS dashboard, this is not always a straightforward process. FAERS writes all the reactions associated with an agent in a single column, such that one column may have over 10-15 reactions. As such, this step will be performed here.*

*The cleaning process has not yet been started. The appropriate R code will be provided as the analysis is carried out.*

## Statistical analysis

*The first of the analysis will consists of descriptive statistics. Continous variables will be summarized as means, whereas categorical variables will be summarized as counts and percenatges.*

*The second part of the analysis will involve a disproportionality analysis. This analysis uses the following formula to calculate the reported odds ratio (ROR).*

*a = irAES reported with ICIs* *b = irAEs reported with all other drugs* *c = All other AEs reported with ICIs* *d = All other AEs reported with other drugs*

*ROR = a/c / b/d*

*Other details in the analysis will be finalized later.*

{{< pagebreak >}}

# Results

*This part has not been completed. The source code has been left as is*

## Exploratory/Descriptive analysis

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References

### 2025-02-21 code/eda-code/eda.qmd
Part 2 of MADA Project

---
title: "Exploratory analysis"
---

We start of by summarizing all the information

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyr)
library(knitr)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.
First, we summarize all our information using a table.

```{r}
library(dplyr)
library(tidyr)
library(kableExtra)

# Create a table for Age summary with Mean and SD
age_summary <- data.frame(
    Category = "Age",
    Value = paste(round(mean(mydata$patient_age, na.rm = TRUE), 2), "±", round(sd(mydata$patient_age, na.rm = TRUE), 2)),
    Count = NA,
    Percent = NA
)

# Sex Summary
sex_summary <- mydata %>%
    group_by(sex) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = sex) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Drug Summary
drug_summary <- mydata %>%
    group_by(suspect_product_active_ingredients) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = suspect_product_active_ingredients) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Outcome Summary
outcome_summary <- mydata %>%
    group_by(serious) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = serious) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Reporter Type Summary
reporter_summary <- mydata %>%
    group_by(reporter_type) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = reporter_type) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Country Summary
country_summary <- mydata %>%
    group_by(country_where_event_occurred) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = country_where_event_occurred) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Reaction Type Summary
# Define the organ system columns
organ_systems <- c("organ_system_1", "organ_system_2", "organ_system_3", "organ_system_4", "organ_system_5")

summary_data <- mydata %>%
    select(all_of(organ_systems)) %>%
    pivot_longer(cols = everything(), values_to = "organ_system") %>%
    filter(!is.na(organ_system)) %>%
    group_by(organ_system) %>%
    summarise(
        Count = n(),
        Percent = round((n() / nrow(mydata)) * 100, 1)
    ) %>%
    rename(Category = organ_system) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Combine all summaries into one table
final_summary_table <- bind_rows(
    data.frame(Category = "Age", Value = age_summary$Value, stringsAsFactors = FALSE),
    data.frame(Category = "Sex", Value = "", stringsAsFactors = FALSE), # Subheading for Sex
    sex_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Sex
    data.frame(Category = "Immune Checkpoint Inhibitors", Value = "", stringsAsFactors = FALSE), # Subheading for Drugs
    drug_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Drugs
    data.frame(Category = "Outcome", Value = "", stringsAsFactors = FALSE), # Subheading for Outcome
    outcome_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Outcome
    data.frame(Category = "Reporter Type", Value = "", stringsAsFactors = FALSE), # Subheading for Reporter Type
    reporter_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Reporter Type
    data.frame(Category = "Country", Value = "", stringsAsFactors = FALSE), # Subheading for Country
    country_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Country
    data.frame(Category = "Reaction Type", Value = "", stringsAsFactors = FALSE), # Subheading for Reaction Type
    summary_data %>% select(Category, Value)
)

# Create a publication-quality table using kableExtra
final_summary_table %>%
    kable("html", caption = "Table 1: Descriptive Statistics", col.names = c("Category", "Value")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
    column_spec(1, bold = TRUE) %>%
    row_spec(which(final_summary_table$Category %in% c("Age", "Sex", "Immune Checkpoint Inhibitors", "Outcome", "Reporter Type", "Country", "Reaction Type")), bold = TRUE, background = "#f7f7f7") %>%
    footnote(general = "This table summarizes the patient data by various categories.")

```

We save the table.

```{r}
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(final_summary_table, file = summarytable_file)
```

# Data exploration through figures

Bar graph for outcomes by drugs.

```{r}
library(ggplot2)

# Define custom colors for the 'serious' variable
custom_colors <- c(
    "Serious" = "#d7191c",  # Red for serious
    "Non-Serious" = "#2c7bb6"    # Blue for non-serious
)

# Create the plot with custom colors for 'serious'
p1 <- mydata %>%
    ggplot(aes(x = suspect_product_active_ingredients, fill = serious)) +
    geom_bar(position = "stack") + # Use "stack" for stacked bars
    scale_fill_manual(values = custom_colors) + # Apply custom colors
    labs(
        title = "Bar Plot of Outcomes of Immune Checkpoint Inhibitors",
        x = "Immune Checkpoint Inhibitors",
        y = "Count",
        fill = "Outcomes"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability

# Print the plot
print(p1)

figure_file = here("results", "figures","figure1.png")
ggsave(filename = figure_file, plot=p1) 
```

Now let's look at the bar graph by type of organ system.

```{r}
p2 <- summary_data %>%
    ggplot(aes(x = reorder(Category, Count), y = Count)) + # Reorder by Count ascending
    geom_col(fill = "#1f77b4") +
    labs(
        title = "Bar Plot of Reaction Types",
        x = "Organ System",
        y = "Count",
        caption = "Percentages are shown in parentheses"
    ) +
    theme_minimal() +
    geom_text(aes(label = Value), hjust = -0.1, size = 3) + # Adjust label position
    coord_flip() # Flip the axes

# Print the plot
print(p2)

figure_file = here("results", "figures","figure2.png")
ggsave(filename = figure_file, plot=p2) 
```

### 2025-02-21 code/processing-code/processingfile-v1.qmd
Part 2 of MADA Project

---
title: "Processing raw data from FAERS"
output: html_document
---

# Processing script

This quarto file has the code to process the raw data.

# Setup

Load needed packages. make sure they are installed.

```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

# Data loading

Loading the data into R

```{r}
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","ICI_FAERS.xlsx")
ici_rawdata <- readxl::read_excel(data_location)
```

# Check data

First we look at the data

```{r}
glimpse(ici_rawdata)
summary(ici_rawdata)
```

# Cleaning the data

First, let's rename all the columns to make them easier to use in R.

```{r}
library(janitor)
ici_rawdata = ici_rawdata %>%
  clean_names()
```

Since our focus is on immune checkpoint inhibitors, we remove any cases where concomitant products are available.

```{r}
ici_rawdata = ici_rawdata %>%
  filter(concomitant_product_names == "-")
```

We now have 1860 cases.

The next step is to remove cases where other drugs are also present in the suspect product active ingredients.

```{r}
library(stringr)
#First create a vector of valid drugs
valid_drugs = c("Atezolizumab", "Nivolumab", "Pembrolizumab", "Avelumab")

#Filter the data
ici_data1 = ici_rawdata %>%
  filter(
    tolower(suspect_product_active_ingredients) %in% tolower(valid_drugs)
  )
```

There are now 1398 cases in the data.

The next step is to filter the data to retain only immune-related adverse events (irAEs)

Now, we will filter the reactions - and retain only irAEs.
Common irAEs have been included - sourced from 2 separate reviews. The details are provided in the methods section.

```{r}
# Create a vector of specific reactions you want to match
reactions_of_interest <- c(
  "colitis", "hyperthyroidism", "hypothyroidism", "hepatitis", "pruritus", "rash", 
  "pneumonitis", "myocarditis", "pericarditis", "dermatitis", "Stevens-Johnson syndrome", 
  "sjogren's syndrome", "toxic epidermal necrolysis", "hypophysitis", "autoimmune diabetes", 
  "pancreatitis", "myositis", "uveitis", "episcleritis", "nephritis", "peripheral neuropathy", "myelitis", "encephalitis", "erythema", "autoimmune", "psoriasis", "adrenal", "adrenitis", "hypercortisolism", "hypocortisolism" ,"hyperparathyroidism", "hypoparathyroidism", "hypogonadism", "thyroiditis", "graves", "thyrotoxicosis", "enterocolitis", "ileitis", "ileocolitis", "myasthenia", "neuritis", "arthritis", "lupus", "sarcoidosis", "vasculitis", "immune"
)

ici_data_subset <- ici_data1 %>%
  filter(str_detect(tolower(reactions), paste(tolower(reactions_of_interest), collapse = "|")))

# View the subset
View(ici_data_subset)
```

We have 368 reactions.
The next step is to characterize and categorize the irAEs.
The first step is to split the reaction column, and keep only the irAEs.

```{r}
library(tidyr)

ici_data_split <- ici_data_subset %>%
  separate(reactions, into = paste0("reaction_", 1:15), sep = ";", remove = FALSE, extra = "drop")

# View the result
View(ici_data_split)

reaction_columns <- paste0("reaction_", 1:15)

##Only retain irAEs
ici_data_split <- ici_data_split %>%
  mutate(across(all_of(reaction_columns), 
                ~ if_else(str_detect(tolower(.), paste(gsub("'", "\\\\'", reactions_of_interest), collapse = "|")), 
                          ., NA_character_)))

```

Now, we combine the columns and remove the NAs.
```{r}
# Define columns to shift
cols_to_shift <- paste0("reaction_", 1:15)
cols_to_keep <- setdiff(names(ici_data_split), cols_to_shift)

# Function to shift non-NA values left within each row
shift_non_na <- function(row) {
  non_na_values <- na.omit(row)  # Get non-NA values
  c(non_na_values, rep(NA, length(row) - length(non_na_values)))  # Fill rest with NA
}

# Apply function only to reaction_* columns
ici_data_split[cols_to_shift] <- t(apply(ici_data_split[cols_to_shift], 1, shift_non_na))

# Remove columns where all variables are NA
ici_data_cleaned = ici_data_split %>% select(where( ~ !all(is.na(.))))
```

The next step is to code these reactions using MedDRA Preferred Terms (PTs), which is a distinct descriptor (single medical concept) for a symptom, sign, disease diagnosis, therapeutic indication, investigation, surgical or medical procedure, and medical social or family history characteristic.
After going through the list of reactions, I found that the reactions are already in PTs.

Now, the next step is to remove columns that are not required.

```{r}
ici_data_final = ici_data_cleaned %>%
  select(-reactions)
```

Now, we convert the patient age data to numeric.

```{r}
ici_data_final <- ici_data_final %>%
  mutate(patient_age = as.numeric(gsub("[^0-9]", "", patient_age)))
```

Let us classify the countries into US and Outside US
```{r}
ici_data_final <- ici_data_final %>%
  mutate(country_where_event_occurred = ifelse(country_where_event_occurred != "US", "Outside US", country_where_event_occurred))

```

```{r}
library(dplyr)

# Define lists of reaction keywords for each organ system
dermatological_reactions <- c("rash", "pruritus", "dermatitis", "psoriasis")
gastrointestinal_reactions <- c("colitis", "hepatitis", "pancreatitis", "enterocolitis", "ileitis", "ileocolitis")
neurological_reactions <- c("myasthenia", "graves", "myelitis", "encephalitis", "neuritis", "peripheral neuropathy")
endocrinological_reactions <- c("hyperthyroidism", "hypothyroidism", "thyroiditis", "thyrotoxicosis", "adrenitis", "adrenal", "hypogonadism", "hyperparathyroidism", "hypercortisolism", "hypophysitis", "hypoparathyroidism", "hypocortisolism", "autoimmune diabetes")
cardiovascular_reactions <- c("myocarditis", "pericarditis")

# Assuming your data is stored in 'ici_data_final' and reactions are in reaction_1 to reaction_5
ici_data_final <- ici_data_final %>%
  mutate(
    organ_system_1 = case_when(
      !is.na(reaction_1) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Skin",
      !is.na(reaction_1) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Gastrointestinal",
      !is.na(reaction_1) & grepl(paste(neurological_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Neurological",
      !is.na(reaction_1) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Endocrinological",
      !is.na(reaction_1) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Cardiovascular",
      !is.na(reaction_1) ~ "Other",  # If not NA and no match, classify as Other
      TRUE ~ NA_character_  # If NA, keep as NA
    ),

    organ_system_2 = case_when(
      !is.na(reaction_2) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Skin",
      !is.na(reaction_2) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Gastrointestinal",
      !is.na(reaction_2) & grepl(paste(neurological_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Neurological",
      !is.na(reaction_2) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Endocrinological",
      !is.na(reaction_2) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Cardiovascular",
      !is.na(reaction_2) ~ "Other",
      TRUE ~ NA_character_
    ),

    organ_system_3 = case_when(
      !is.na(reaction_3) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Skin",
      !is.na(reaction_3) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Gastrointestinal",
      !is.na(reaction_3) & grepl(paste(neurological_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Neurological",
      !is.na(reaction_3) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Endocrinological",
      !is.na(reaction_3) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Cardiovascular",
      !is.na(reaction_3) ~ "Other",
      TRUE ~ NA_character_
    ),

    organ_system_4 = case_when(
      !is.na(reaction_4) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Skin",
      !is.na(reaction_4) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Gastrointestinal",
      !is.na(reaction_4) & grepl(paste(neurological_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Neurological",
      !is.na(reaction_4) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Endocrinological",
      !is.na(reaction_4) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Cardiovascular",
      !is.na(reaction_4) ~ "Other",
      TRUE ~ NA_character_
    ),

    organ_system_5 = case_when(
      !is.na(reaction_5) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Skin",
      !is.na(reaction_5) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Gastrointestinal",
      !is.na(reaction_5) & grepl(paste(neurological_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Neurological",
      !is.na(reaction_5) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Endocrinological",
      !is.na(reaction_5) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Cardiovascular",
      !is.na(reaction_5) ~ "Other",
      TRUE ~ NA_character_
    )
  )

# Check the new columns
head(ici_data_final)

```

This is our final data.
Next, we move on to exploratory data analysis.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata = ici_data_final
```

# Save data

Finally, we save the clean data as RDS file.

```{r}
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

### 2025-02-21 products/manuscript/Manuscript.qmd
Part 2 of MADA Project

@@
 
-*Write a summary of your project.*
-*This part has not been completed. The source code has been left as is*
+*Write a summary of your project.* *This part has not been completed. The source code has been left as is*
 
@@
 
-*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*
-*This part has not been completed. The source code has been left as is*
-*Details have been provided in the data acquisition and statistical analysis parts.*
-
-## Schematic of workflow
-
-*This part has not been completed. The source code has been left as is*
-
-Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-@fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
-We store those figures in the `assets` folder.
-
-```{r}
-#| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
-#| echo: FALSE
-knitr::include_graphics(here("assets","antigen-recognition.png"))
-```
+*Describe your methods.
+That should describe the data, the cleaning processes, and the analysis approaches.
+You might want to provide a shorter description here and all the details in the supplement.* *This part has not been completed.
+The source code has been left as is* *Details have been provided in the data acquisition and statistical analysis parts.*
 
@@
 
-*The cleaning process has not yet been started. The appropriate R code will be provided as the analysis is carried out.*
+*The FAERS data initially consisted of 2948 reactions associated with ICIs in bladder cancer.*
+
+*Reactions with concomitant suspect products, combination products, and those which were not irAEs were removed.
+This resulted in a total of 369 cases.
+The details are available in the processing code file.*
 
@@
 
-*Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project.*
+@tbl-summarytable *provides the patient characteristics, and summary of reactions by type of immune checkpoint inhibitor, and organ system affected.*
 
-@tbl-summarytable shows a summary of the data.
+*The mean age reported was 66.94±17.65. A large percentage of the reactions occurred in males (70.2%). This may be because bladder cancer is much more common in males than in females.*
 
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-I generally recommend the `here` package.
+*Over 60% of the reactions were associated with pembrolizumab. A majority of the reactions were serious (83.7%)*
 
@@
 #| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
+#| tbl-cap: "Table 1: Descriptive Statistics"
 #| echo: FALSE
@@
 
-*To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
+*@fig-result shows the number of reactions by drug, along with the seriousness of the reaction.*
 
@@
 #| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
+#| fig-cap: "Bar chart of outcome of immune checkpoint inhibitors"
 #| echo: FALSE
-knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
+knitr::include_graphics(here("results","figures","figure1.png"))
+```
+
+*@fig-result1 shows the bar chart by type of organ system involved in the reaction.*
+
+```{r}
+#| label: fig-result1
+#| fig-cap: "Bar chart of types of reactions"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","figure2.png"))
 ```
@@
 ## Full analysis
+
+*This part has not yet been edited.*
 

### 2025-03-14 code/analysis-code/statistical-analysis.qmd
Part 3 of MADA Project

---
title: "statistical-analysis"
format: html
---

This file contains details of the statistical analysis.

After the exploratory data analysis and the descriptive statistics, the next step is to run a disproportionality analysis. The first step is to assess the adverse events in comparison to the entire FAERS database.

To assess this, we need the following information: 
a = Number of irAEs with 1 ICI (e.g. atezolizumab) 
b = Number of reactions other than irAEs with atezolizumab
c = Number of irAEs with drugs other than atezolizumab 
d = Number of reactions other than irAEs with drugs other than atzeolizumab.

The same procedure would be repeated for the other 3 ICIs.

First, we load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

First we find number of irAEs with each ICI.

```{r}
# Load necessary library
library(dplyr)

# List of drugs to analyze
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Function to count total reactions for a given drug
count_total_reactions <- function(data, drug) {
  # Filter rows for the drug
  drug_data <- data %>%
    filter(grepl(drug, suspect_product_active_ingredients, ignore.case = TRUE))

  # Combine reaction columns into a single vector
  all_reactions <- c(
    drug_data$reaction_1,
    drug_data$reaction_2,
    drug_data$reaction_3,
    drug_data$reaction_4,
    drug_data$reaction_5
  )

  # Remove NA values
  all_reactions <- all_reactions[!is.na(all_reactions)]

  # Count total reactions
  total_reactions <- length(all_reactions)

  # Return total reactions
  return(total_reactions)
}

# Count total reactions for each drug
results <- sapply(drugs, function(drug) count_total_reactions(mydata, drug))

# Print the results
results_irae <- data.frame(Drug = drugs, Total_Reactions = results)
print(results_irae)
```

The `echo: false` option disables the printing of code (only output is displayed).

Now, we do the same for reactions other than irAEs. We created and saved this data when we processed it.

Load the data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata1.rds")
#load data
mydata1 <- readRDS(data_location)
```

Now we find number of reactions other than irAEs with each ICI.

```{r}
# Load necessary library
library(dplyr)

# List of drugs to analyze
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Function to count total reactions for a given drug
count_total_reactions1 <- function(data, drug) {
  # Filter rows for the drug
  drug_data1 <- data %>%
    filter(grepl(drug, suspect_product_active_ingredients, ignore.case = TRUE))

  # Combine reaction columns into a single vector
  all_reactions1 <- c(
    drug_data1$reaction_1,
    drug_data1$reaction_2,
    drug_data1$reaction_3,
    drug_data1$reaction_4,
    drug_data1$reaction_5
  )

  # Remove NA values
  all_reactions1 <- all_reactions1[!is.na(all_reactions1)]

  # Count total reactions
  total_reactions1 <- length(all_reactions1)

  # Return total reactions
  return(total_reactions1)
}

# Count total reactions for each drug
results1 <- sapply(drugs, function(drug) count_total_reactions1(mydata1, drug))

# Print the results
results_non_irae <- data.frame(Drug = drugs, Total_Reactions = results1)
print(results_non_irae)
```
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for All irAEs
results_table_all_iraes <- data.frame()

# Loop through each drug and perform the analysis for all irAEs
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for all irAEs combined)
  a <- results_irae %>% filter(Drug == drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()
  b <- results_non_irae %>% filter(Drug == drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()
  c <- results_irae %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()
  d <- results_non_irae %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a / (a + b)) / (c / (c + d))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a) - (1 / (a + b)) + (1 / c) - (1 / (c + d)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a / b) / (c / d)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a) + (1 / b) + (1 / c) + (1 / d))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a * (a + b + c + d)) / ((a + c) * (a + b)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a - 1 / (a + b) - 1 / (a + c) + 1 / (a + b + c + d))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (All irAEs)
  results_table_all_iraes <- rbind(results_table_all_iraes, data.frame(
    Drug = drug_of_interest,
    Reaction = "All irAEs",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for All irAEs
print(results_table_all_iraes)
```
The next step is to assess the safety signal for individual irAEs.
The commonly reported irAEs in clinical trials are rash, pruritus, colitis, hepatitis, hyperthyroidism, hypothyroidism, and nephritis.
We conduct disproportionality analysis for each of these.

We compare each of these reactions to all other reactions.
First, we assess the number of 'rash' reactions reported with all the ICIs.

```{r}
 # Filter for the specific drugs and count 'Rash' reactions
rash_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Rash = sum(c_across(starts_with("reaction_")) == "Rash", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Rash_Count = sum(Rash))

# Print the result
print(rash_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'rash' reactions associated with each ICI.

```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

rash_counts <- rash_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'Rash' counts from total reactions
results_non_rash <- results_irae %>%
  mutate(
    Non_Rash_Reactions = Total_Reactions - rash_counts$Rash_Count
  )

# Print the updated results
print(results_non_rash)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_rash <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_rash$Non_Rash_Reactions
)

# Print the new dataset
print(results_other_rash)

```

We have all our numbers now.

```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Rash
results_table_rash <- data.frame()

# Loop through each drug and perform the analysis for Rash
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Rash)
  a_rash <- rash_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Rash_Count)
  b_rash <- results_other_rash %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_rash <- rash_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Rash_Count)) %>% pull()
  d_rash <- results_other_rash %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_rash / (a_rash + b_rash)) / (c_rash / (c_rash + d_rash))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_rash) - (1 / (a_rash + b_rash)) + (1 / c_rash) - (1 / (c_rash + d_rash)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_rash / b_rash) / (c_rash / d_rash)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_rash) + (1 / b_rash) + (1 / c_rash) + (1 / d_rash))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_rash * (a_rash + b_rash + c_rash + d_rash)) / ((a_rash + c_rash) * (a_rash + b_rash)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_rash - 1 / (a_rash + b_rash) - 1 / (a_rash + c_rash) + 1 / (a_rash + b_rash + c_rash + d_rash))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Rash)
  results_table_rash <- rbind(results_table_rash, data.frame(
    Drug = drug_of_interest,
    Reaction = "Rash",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Rash
print(results_table_rash)

```

Now, we conduct the same analysis for pruritus.
```{r}
 # Filter for the specific drugs and count 'Pruritus' reactions
pruritus_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Pruritus = sum(c_across(starts_with("reaction_")) == "Pruritus", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Pruritus_Count = sum(Pruritus))

# Print the result
print(pruritus_counts)
```

We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'pruritus' reactions associated with each ICI.

```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

pruritus_counts <- pruritus_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'pruritus' counts from total reactions
results_non_pruritus <- results_irae %>%
  mutate(
    Non_Pruritus_Reactions = Total_Reactions - pruritus_counts$Pruritus_Count
  )

# Print the updated results
print(results_non_pruritus)

# Create the new dataset by adding Non_Pruritus_Reactions to Total_Reactions
results_other_pruritus <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_pruritus$Non_Pruritus_Reactions
)

# Print the new dataset
print(results_other_pruritus)

```
We now conduct the analysis.

```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Pruritus
results_table_pruritus <- data.frame()

# Loop through each drug and perform the analysis for Pruritus
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Pruritus)
  a_pruritus <- pruritus_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Pruritus_Count)
  b_pruritus <- results_other_pruritus %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_pruritus <- pruritus_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Pruritus_Count)) %>% pull()
  d_pruritus <- results_other_pruritus %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_pruritus / (a_pruritus + b_pruritus)) / (c_pruritus / (c_pruritus + d_pruritus))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_pruritus) - (1 / (a_pruritus + b_pruritus)) + (1 / c_pruritus) - (1 / (c_pruritus + d_pruritus)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_pruritus / b_pruritus) / (c_pruritus / d_pruritus)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_pruritus) + (1 / b_pruritus) + (1 / c_pruritus) + (1 / d_pruritus))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_pruritus * (a_pruritus + b_pruritus + c_pruritus + d_pruritus)) / ((a_pruritus + c_pruritus) * (a_pruritus + b_pruritus)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_pruritus - 1 / (a_pruritus + b_pruritus) - 1 / (a_pruritus + c_pruritus) + 1 / (a_pruritus + b_pruritus + c_pruritus + d_pruritus))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Pruritus)
  results_table_pruritus <- rbind(results_table_pruritus, data.frame(
    Drug = drug_of_interest,
    Reaction = "Pruritus",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Pruritus
print(results_table_pruritus)

```

We repeat this analysis for hypothyroidism.
```{r}
 # Filter for the specific drugs and count 'Hypothyroidism' reactions
hypothyroidism_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Hypothyroidism = sum(c_across(starts_with("reaction_")) == "Hypothyroidism", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Hypothyroidism_Count = sum(Hypothyroidism))

# Print the result
print(hypothyroidism_counts)
```

We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'hypothyroidism' reactions associated with each ICI.

```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

hypothyroidism_counts <- hypothyroidism_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'Hypothyroidism' counts from total reactions
results_non_hypothyroidism <- results_irae %>%
  mutate(
    Non_Hypothyroidism_Reactions = Total_Reactions - hypothyroidism_counts$Hypothyroidism_Count
  )

# Print the updated results
print(results_non_hypothyroidism)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_hypothyroidism <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_hypothyroidism$Non_Hypothyroidism_Reactions
)

# Print the new dataset
print(results_other_hypothyroidism)

```

Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Hypothyroidism
results_table_hypothyroidism <- data.frame()

# Loop through each drug and perform the analysis for Hypothyroidism
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Hypothyroidism)
  a_hypothyroidism <- hypothyroidism_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Hypothyroidism_Count)
  b_hypothyroidism <- results_other_hypothyroidism %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_hypothyroidism <- hypothyroidism_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Hypothyroidism_Count)) %>% pull()
  d_hypothyroidism <- results_other_hypothyroidism %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_hypothyroidism / (a_hypothyroidism + b_hypothyroidism)) / (c_hypothyroidism / (c_hypothyroidism + d_hypothyroidism))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_hypothyroidism) - (1 / (a_hypothyroidism + b_hypothyroidism)) + (1 / c_hypothyroidism) - (1 / (c_hypothyroidism + d_hypothyroidism)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_hypothyroidism / b_hypothyroidism) / (c_hypothyroidism / d_hypothyroidism)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_hypothyroidism) + (1 / b_hypothyroidism) + (1 / c_hypothyroidism) + (1 / d_hypothyroidism))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_hypothyroidism * (a_hypothyroidism + b_hypothyroidism + c_hypothyroidism + d_hypothyroidism)) / ((a_hypothyroidism + c_hypothyroidism) * (a_hypothyroidism + b_hypothyroidism)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_hypothyroidism - 1 / (a_hypothyroidism + b_hypothyroidism) - 1 / (a_hypothyroidism + c_hypothyroidism) + 1 / (a_hypothyroidism + b_hypothyroidism + c_hypothyroidism + d_hypothyroidism))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Hypothyroidism)
  results_table_hypothyroidism <- rbind(results_table_hypothyroidism, data.frame(
    Drug = drug_of_interest,
    Reaction = "Hypothyroidism",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Hypothyroidism
print(results_table_hypothyroidism)

```

Now, we assess the reports for hyperthyroidism:
```{r}
 # Filter for the specific drugs and count 'hyperthyroidism' reactions
hyperthyroidism_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Hyperthyroidism = sum(c_across(starts_with("reaction_")) == "Hyperthyroidism", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Hyperthyroidism_Count = sum(Hyperthyroidism))

# Print the result
print(hyperthyroidism_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'hyperthyroidism' reactions associated with each ICI.
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

hyperthyroidism_counts <- hyperthyroidism_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'Hyperthyroidism' counts from total reactions
results_non_hyperthyroidism <- results_irae %>%
  mutate(
    Non_Hyperthyroidism_Reactions = Total_Reactions - hyperthyroidism_counts$Hyperthyroidism_Count
  )

# Print the updated results
print(results_non_hyperthyroidism)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_hyperthyroidism <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_hyperthyroidism$Non_Hyperthyroidism_Reactions
)

# Print the new dataset
print(results_other_hyperthyroidism)
```

Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Hyperthyroidism
results_table_hyperthyroidism <- data.frame()

# Loop through each drug and perform the analysis for Hyperthyroidism
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Hyperthyroidism)
  a_hyperthyroidism <- hyperthyroidism_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Hyperthyroidism_Count)
  b_hyperthyroidism <- results_other_hyperthyroidism %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_hyperthyroidism <- hyperthyroidism_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Hyperthyroidism_Count)) %>% pull()
  d_hyperthyroidism <- results_other_hyperthyroidism %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_hyperthyroidism / (a_hyperthyroidism + b_hyperthyroidism)) / (c_hyperthyroidism / (c_hyperthyroidism + d_hyperthyroidism))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_hyperthyroidism) - (1 / (a_hyperthyroidism + b_hyperthyroidism)) + (1 / c_hyperthyroidism) - (1 / (c_hyperthyroidism + d_hyperthyroidism)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_hyperthyroidism / b_hyperthyroidism) / (c_hyperthyroidism / d_hyperthyroidism)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_hyperthyroidism) + (1 / b_hyperthyroidism) + (1 / c_hyperthyroidism) + (1 / d_hyperthyroidism))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_hyperthyroidism * (a_hyperthyroidism + b_hyperthyroidism + c_hyperthyroidism + d_hyperthyroidism)) / ((a_hyperthyroidism + c_hyperthyroidism) * (a_hyperthyroidism + b_hyperthyroidism)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_hyperthyroidism - 1 / (a_hyperthyroidism + b_hyperthyroidism) - 1 / (a_hyperthyroidism + c_hyperthyroidism) + 1 / (a_hyperthyroidism + b_hyperthyroidism + c_hyperthyroidism + d_hyperthyroidism))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Hyperthyroidism)
  results_table_hyperthyroidism <- rbind(results_table_hyperthyroidism, data.frame(
    Drug = drug_of_interest,
    Reaction = "Hyperthyroidism",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Hyperthyroidism
print(results_table_hyperthyroidism)

```

Now, we repeat this analysis with colitis:

```{r}
 # Filter for the specific drugs and count 'colitis' reactions
colitis_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Colitis = sum(c_across(starts_with("reaction_")) == "Colitis", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Colitis_Count = sum(Colitis))

# Print the result
print(colitis_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'colitis' reactions associated with each ICI.
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

colitis_counts <- colitis_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'colitis' counts from total reactions
results_non_colitis <- results_irae %>%
  mutate(
    Non_Colitis_Reactions = Total_Reactions - colitis_counts$Colitis_Count
  )

# Print the updated results
print(results_non_colitis)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_colitis <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_colitis$Non_Colitis_Reactions
)

# Print the new dataset
print(results_other_colitis)
```

Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Colitis
results_table_colitis <- data.frame()

# Loop through each drug and perform the analysis for Colitis
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Colitis)
  a_colitis <- colitis_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Colitis_Count)
  b_colitis <- results_other_colitis %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_colitis <- colitis_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Colitis_Count)) %>% pull()
  d_colitis <- results_other_colitis %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_colitis / (a_colitis + b_colitis)) / (c_colitis / (c_colitis + d_colitis))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_colitis) - (1 / (a_colitis + b_colitis)) + (1 / c_colitis) - (1 / (c_colitis + d_colitis)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_colitis / b_colitis) / (c_colitis / d_colitis)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_colitis) + (1 / b_colitis) + (1 / c_colitis) + (1 / d_colitis))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_colitis * (a_colitis + b_colitis + c_colitis + d_colitis)) / ((a_colitis + c_colitis) * (a_colitis + b_colitis)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_colitis - 1 / (a_colitis + b_colitis) - 1 / (a_colitis + c_colitis) + 1 / (a_colitis + b_colitis + c_colitis + d_colitis))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Colitis)
  results_table_colitis <- rbind(results_table_colitis, data.frame(
    Drug = drug_of_interest,
    Reaction = "Colitis",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Colitis
print(results_table_colitis)

```
Now, we conduct the same analysis for hepatitis:

```{r}
# Filter for the specific drugs and count 'hepatitis' reactions
hepatitis_counts <- mydata %>%
  filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
  rowwise() %>%
  mutate(Hepatitis = sum(c_across(starts_with("reaction_")) == "Hepatitis", na.rm = TRUE)) %>%
  group_by(suspect_product_active_ingredients) %>%
  summarise(Hepatitis_Count = sum(Hepatitis))

# Print the result
print(hepatitis_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'hepatitis' reactions associated with each ICI.
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

hepatitis_counts <- hepatitis_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'hepatitis' counts from total reactions
results_non_hepatitis <- results_irae %>%
  mutate(
    Non_Hepatitis_Reactions = Total_Reactions - hepatitis_counts$Hepatitis_Count
  )

# Print the updated results
print(results_non_hepatitis)

# Create the new dataset by adding Non_Hepatitis_Reactions to Total_Reactions
results_other_hepatitis <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_hepatitis$Non_Hepatitis_Reactions
)

# Print the new dataset
print(results_other_hepatitis)
```
Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Hepatitis
results_table_hepatitis <- data.frame()

# Loop through each drug and perform the analysis for Hepatitis
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Hepatitis)
  a_hepatitis <- hepatitis_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Hepatitis_Count)
  b_hepatitis <- results_other_hepatitis %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_hepatitis <- hepatitis_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Hepatitis_Count)) %>% pull()
  d_hepatitis <- results_other_hepatitis %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_hepatitis / (a_hepatitis + b_hepatitis)) / (c_hepatitis / (c_hepatitis + d_hepatitis))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_hepatitis) - (1 / (a_hepatitis + b_hepatitis)) + (1 / c_hepatitis) - (1 / (c_hepatitis + d_hepatitis)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_hepatitis / b_hepatitis) / (c_hepatitis / d_hepatitis)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_hepatitis) + (1 / b_hepatitis) + (1 / c_hepatitis) + (1 / d_hepatitis))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_hepatitis * (a_hepatitis + b_hepatitis + c_hepatitis + d_hepatitis)) / ((a_hepatitis + c_hepatitis) * (a_hepatitis + b_hepatitis)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_hepatitis - 1 / (a_hepatitis + b_hepatitis) - 1 / (a_hepatitis + c_hepatitis) + 1 / (a_hepatitis + b_hepatitis + c_hepatitis + d_hepatitis))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Hepatitis)
  results_table_hepatitis <- rbind(results_table_hepatitis, data.frame(
    Drug = drug_of_interest,
    Reaction = "Hepatitis",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Hepatitis
print(results_table_hepatitis)

```
We repeat this with nephritis:
```{r}
# Filter for the specific drugs and count 'nephritis' reactions
nephritis_counts <- mydata %>%
  filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
  rowwise() %>%
  mutate(Nephritis = sum(c_across(starts_with("reaction_")) == "Nephritis", na.rm = TRUE)) %>%
  group_by(suspect_product_active_ingredients) %>%
  summarise(Nephritis_Count = sum(Nephritis))

# Print the result
print(nephritis_counts)
```
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

nephritis_counts <- nephritis_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'nephritis' counts from total reactions
results_non_nephritis <- results_irae %>%
  mutate(
    Non_Nephritis_Reactions = Total_Reactions - nephritis_counts$Nephritis_Count
  )

# Print the updated results
print(results_non_nephritis)

# Create the new dataset by adding Non_Nephritis_Reactions to Total_Reactions
results_other_nephritis <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_nephritis$Non_Nephritis_Reactions
)

# Print the new dataset
print(results_other_nephritis)
```
Disproportionality analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Nephritis
results_table_nephritis <- data.frame()

# Loop through each drug and perform the analysis for Nephritis
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Nephritis)
  a_nephritis <- nephritis_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Nephritis_Count)
  b_nephritis <- results_other_nephritis %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_nephritis <- nephritis_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Nephritis_Count)) %>% pull()
  d_nephritis <- results_other_nephritis %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_nephritis / (a_nephritis + b_nephritis)) / (c_nephritis / (c_nephritis + d_nephritis))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_nephritis) - (1 / (a_nephritis + b_nephritis)) + (1 / c_nephritis) - (1 / (c_nephritis + d_nephritis)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_nephritis / b_nephritis) / (c_nephritis / d_nephritis)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_nephritis) + (1 / b_nephritis) + (1 / c_nephritis) + (1 / d_nephritis))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_nephritis * (a_nephritis + b_nephritis + c_nephritis + d_nephritis)) / ((a_nephritis + c_nephritis) * (a_nephritis + b_nephritis)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_nephritis - 1 / (a_nephritis + b_nephritis) - 1 / (a_nephritis + c_nephritis) + 1 / (a_nephritis + b_nephritis + c_nephritis + d_nephritis))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Nephritis)
  results_table_nephritis <- rbind(results_table_nephritis, data.frame(
    Drug = drug_of_interest,
    Reaction = "Nephritis",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Nephritis
print(results_table_nephritis)

```
```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(huxtable)

# Combine all individual results tables
all_results <- bind_rows(
  results_table_all_iraes,
  results_table_rash,
  results_table_pruritus,
  results_table_hypothyroidism,
  results_table_hyperthyroidism,
  results_table_colitis,
  results_table_hepatitis,
  results_table_nephritis
)

# Pivot data to get reactions as rows and drugs as grouped columns with subcolumns for PRR, ROR, and IC
summary_table <- all_results %>%
  pivot_wider(names_from = Drug, values_from = c(PRR, ROR, IC),
              names_glue = "{Drug} {.value} (95% CI)")

# Check the structure of the table and columns
colnames(summary_table)

# Highlight significant values using conditional bolding (PRR > 1, ROR > 1, IC > 0)
summary_table <- summary_table %>%
  mutate(across(starts_with("PRR"), ~ ifelse(. > 1 & grepl("\\(1", .), "bold", .))) %>%
  mutate(across(starts_with("ROR"), ~ ifelse(. > 1 & grepl("\\(1", .), "bold", .))) %>%
  mutate(across(starts_with("IC"), ~ ifelse(. > 0 & grepl("\\(0", .), "bold", .)))

# Split the summary_table into 4 separate tables for each drug
# Table for Atezolizumab
table_atezolizumab <- summary_table %>%
  select(Reaction, starts_with("Atezolizumab")) %>%
  as_hux() %>%
  set_bold(1, 1:ncol(.)) %>%
  set_caption("Atezolizumab Drug-Reaction Analysis") %>%
  set_width(1)

# Apply bolding for significant values for Atezolizumab
for (i in 2:ncol(table_atezolizumab)) {
  table_atezolizumab <- table_atezolizumab %>%
    set_bold(which(table_atezolizumab[, i] == "bold"), i)
}

# Table for Avelumab
table_avelumab <- summary_table %>%
  select(Reaction, starts_with("Avelumab")) %>%
  as_hux() %>%
  set_bold(1, 1:ncol(.)) %>%
  set_caption("Avelumab Drug-Reaction Analysis") %>%
  set_width(1)

# Apply bolding for significant values for Avelumab
for (i in 2:ncol(table_avelumab)) {
  table_avelumab <- table_avelumab %>%
    set_bold(which(table_avelumab[, i] == "bold"), i)
}

# Table for Nivolumab
table_nivolumab <- summary_table %>%
  select(Reaction, starts_with("Nivolumab")) %>%
  as_hux() %>%
  set_bold(1, 1:ncol(.)) %>%
  set_caption("Nivolumab Drug-Reaction Analysis") %>%
  set_width(1)

# Apply bolding for significant values for Nivolumab
for (i in 2:ncol(table_nivolumab)) {
  table_nivolumab <- table_nivolumab %>%
    set_bold(which(table_nivolumab[, i] == "bold"), i)
}

# Table for Pembrolizumab
table_pembrolizumab <- summary_table %>%
  select(Reaction, starts_with("Pembrolizumab")) %>%
  as_hux() %>%
  set_bold(1, 1:ncol(.)) %>%
  set_caption("Pembrolizumab Drug-Reaction Analysis") %>%
  set_width(1)

# Apply bolding for significant values for Pembrolizumab
for (i in 2:ncol(table_pembrolizumab)) {
  table_pembrolizumab <- table_pembrolizumab %>%
    set_bold(which(table_pembrolizumab[, i] == "bold"), i)
}

# Display each table
table_atezolizumab
table_avelumab
table_nivolumab
table_pembrolizumab

```

```{r}
#Save the table
library(here)
# Load necessary libraries
library(here)

# Define paths for saving each table
table_atezolizumab_path <- here("results", "tables", "table_atezolizumab.rds")
table_avelumab_path <- here("results", "tables", "table_avelumab.rds")
table_nivolumab_path <- here("results", "tables", "table_nivolumab.rds")
table_pembrolizumab_path <- here("results", "tables", "table_pembrolizumab.rds")

# Save all four tables to their respective paths
saveRDS(table_atezolizumab, file = table_atezolizumab_path)
saveRDS(table_avelumab, file = table_avelumab_path)
saveRDS(table_nivolumab, file = table_nivolumab_path)
saveRDS(table_pembrolizumab, file = table_pembrolizumab_path)
```

### 2025-03-14 code/eda-code/eda.qmd
Part 3 of MADA Project

@@
 
-Showing a bit of code to produce and save a summary table.
 First, we summarize all our information using a table.
@@
 ```{r}
+# Load necessary libraries
 library(dplyr)
 library(tidyr)
-library(kableExtra)
+library(huxtable)
 
-# Create a table for Age summary with Mean and SD
-age_summary <- data.frame(
-    Category = "Age",
-    Value = paste(round(mean(mydata$patient_age, na.rm = TRUE), 2), "±", round(sd(mydata$patient_age, na.rm = TRUE), 2)),
-    Count = NA,
-    Percent = NA
-)
+# 1. Summarize Age (Mean and SD)
+age_summary <- mydata %>%
+  group_by(suspect_product_active_ingredients) %>%
+  summarise(
+    mean_age = mean(patient_age, na.rm = TRUE),
+    sd_age = sd(patient_age, na.rm = TRUE)
+  ) %>%
+  mutate(`Age, mean years ± SD` = paste0(round(mean_age, 1), " ± ", round(sd_age, 1))) %>%
+  select(suspect_product_active_ingredients, `Age, mean years ± SD`)
 
-# Sex Summary
+# 2. Summarize Sex Distribution
 sex_summary <- mydata %>%
-    group_by(sex) %>%
-    summarise(Count = n()) %>%
-    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
-    rename(Category = sex) %>%
-    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))
+  group_by(suspect_product_active_ingredients, sex) %>%
+  summarise(count = n(), .groups = 'drop') %>%
+  pivot_wider(names_from = sex, values_from = count, values_fill = list(count = 0)) %>%
+  rename(`Not Specified (Sex)` = `Not Specified`) %>%
+  select(suspect_product_active_ingredients, Male, Female, `Not Specified (Sex)`)
 
-# Drug Summary
-drug_summary <- mydata %>%
-    group_by(suspect_product_active_ingredients) %>%
-    summarise(Count = n()) %>%
-    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
-    rename(Category = suspect_product_active_ingredients) %>%
-    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))
+# 3. Summarize Reaction Severity
+reaction_severity_summary <- mydata %>%
+  group_by(suspect_product_active_ingredients, serious) %>%
+  summarise(count = n(), .groups = 'drop') %>%
+  pivot_wider(names_from = serious, values_from = count, values_fill = list(count = 0)) %>%
+  rename(Serious = Serious, `Non-Serious (Reaction)` = `Non-Serious`) %>%
+  select(suspect_product_active_ingredients, Serious, `Non-Serious (Reaction)`)
 
-# Outcome Summary
-outcome_summary <- mydata %>%
-    group_by(serious) %>%
-    summarise(Count = n()) %>%
-    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
-    rename(Category = serious) %>%
-    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))
+# 4. Summarize Outcomes
+outcomes_summary <- mydata %>%
+  group_by(suspect_product_active_ingredients, outcomes) %>%
+  summarise(count = n(), .groups = 'drop') %>%
+  pivot_wider(names_from = outcomes, values_from = count, values_fill = list(count = 0)) %>%
+  rename(`Non-Serious (Outcome)` = `Non-Serious`) %>%
+  select(suspect_product_active_ingredients, Died, Hospitalized, Disabled, `Life Threatening`, `Non-Serious (Outcome)`, `Other Outcomes`)
 
-# Reporter Type Summary
-reporter_summary <- mydata %>%
-    group_by(reporter_type) %>%
-    summarise(Count = n()) %>%
-    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
-    rename(Category = reporter_type) %>%
-    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))
+# 5. Summarize Reporter Type
+reporter_type_summary <- mydata %>%
+  group_by(suspect_product_active_ingredients, reporter_type) %>%
+  summarise(count = n(), .groups = 'drop') %>%
+  pivot_wider(names_from = reporter_type, values_from = count, values_fill = list(count = 0)) %>%
+  select(suspect_product_active_ingredients, `Healthcare Professional`, Consumer)
 
-# Country Summary
+# 6. Summarize Country
 country_summary <- mydata %>%
-    group_by(country_where_event_occurred) %>%
-    summarise(Count = n()) %>%
-    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
-    rename(Category = country_where_event_occurred) %>%
-    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))
-
-# Reaction Type Summary
-# Define the organ system columns
-organ_systems <- c("organ_system_1", "organ_system_2", "organ_system_3", "organ_system_4", "organ_system_5")
-
-summary_data <- mydata %>%
-    select(all_of(organ_systems)) %>%
-    pivot_longer(cols = everything(), values_to = "organ_system") %>%
-    filter(!is.na(organ_system)) %>%
-    group_by(organ_system) %>%
-    summarise(
-        Count = n(),
-        Percent = round((n() / nrow(mydata)) * 100, 1)
-    ) %>%
-    rename(Category = organ_system) %>%
-    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))
+  group_by(suspect_product_active_ingredients, country_where_event_occurred) %>%
+  summarise(count = n(), .groups = 'drop') %>%
+  pivot_wider(names_from = country_where_event_occurred, values_from = count, values_fill = list(count = 0)) %>%
+  rename(`Not Specified (Country)` = `Not Specified`) %>%
+  select(suspect_product_active_ingredients, US, `Outside US`, `Not Specified (Country)`)
 
 # Combine all summaries into one table
-final_summary_table <- bind_rows(
-    data.frame(Category = "Age", Value = age_summary$Value, stringsAsFactors = FALSE),
-    data.frame(Category = "Sex", Value = "", stringsAsFactors = FALSE), # Subheading for Sex
-    sex_summary %>% select(Category, Value),
-    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Sex
-    data.frame(Category = "Immune Checkpoint Inhibitors", Value = "", stringsAsFactors = FALSE), # Subheading for Drugs
-    drug_summary %>% select(Category, Value),
-    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Drugs
-    data.frame(Category = "Outcome", Value = "", stringsAsFactors = FALSE), # Subheading for Outcome
-    outcome_summary %>% select(Category, Value),
-    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Outcome
-    data.frame(Category = "Reporter Type", Value = "", stringsAsFactors = FALSE), # Subheading for Reporter Type
-    reporter_summary %>% select(Category, Value),
-    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Reporter Type
-    data.frame(Category = "Country", Value = "", stringsAsFactors = FALSE), # Subheading for Country
-    country_summary %>% select(Category, Value),
-    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Country
-    data.frame(Category = "Reaction Type", Value = "", stringsAsFactors = FALSE), # Subheading for Reaction Type
-    summary_data %>% select(Category, Value)
-)
+combined_summary <- age_summary %>%
+  full_join(sex_summary, by = "suspect_product_active_ingredients") %>%
+  full_join(reaction_severity_summary, by = "suspect_product_active_ingredients") %>%
+  full_join(outcomes_summary, by = "suspect_product_active_ingredients") %>%
+  full_join(reporter_type_summary, by = "suspect_product_active_ingredients") %>%
+  full_join(country_summary, by = "suspect_product_active_ingredients")
 
-# Create a publication-quality table using kableExtra
-final_summary_table %>%
-    kable("html", caption = "Table 1: Descriptive Statistics", col.names = c("Category", "Value")) %>%
-    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
-    column_spec(1, bold = TRUE) %>%
-    row_spec(which(final_summary_table$Category %in% c("Age", "Sex", "Immune Checkpoint Inhibitors", "Outcome", "Reporter Type", "Country", "Reaction Type")), bold = TRUE, background = "#f7f7f7") %>%
-    footnote(general = "This table summarizes the patient data by various categories.")
+# Convert all columns (except `suspect_product_active_ingredients`) to character
+combined_summary <- combined_summary %>%
+  mutate(across(-suspect_product_active_ingredients, as.character))
+
+# Transpose the table for Word compatibility
+transposed_table <- combined_summary %>%
+  pivot_longer(cols = -suspect_product_active_ingredients, names_to = "Attribute", values_to = "Value") %>%
+  pivot_wider(names_from = suspect_product_active_ingredients, values_from = Value)
+
+# Rename columns for clarity
+colnames(transposed_table)[1] <- "Attribute"
+
+# Convert to huxtable object
+huxtable_table <- as_hux(transposed_table)
+
+# Format the table with huxtable (style, borders, etc.)
+huxtable_table <- huxtable_table %>%
+  set_all_borders(1) %>%
+  set_font_size(12) %>%
+  set_align("center") %>%
+  set_width(1.5)
+
+# Print the huxtable object, this will render correctly in Word
+huxtable_table
 
@@
 # save to file
-summarytable_file = here("results", "tables", "summarytable.rds")
-saveRDS(final_summary_table, file = summarytable_file)
+huxtable_table_path = here("results", "tables", "summarytable.rds")
+saveRDS(huxtable_table, file = huxtable_table_path)
+```
+
+Table 2 - Summarizing organ system reaction by drug type
+```{r}
+# Load necessary libraries
+library(huxtable)
+library(dplyr)
+library(tidyr)
+
+# Define the organ system columns
+organ_systems <- c("organ_system_1", "organ_system_2", "organ_system_3", "organ_system_4", "organ_system_5")
+
+# Transform the organ system columns into a single column and calculate counts and percentages
+organ_system_summary <- mydata %>%
+    select(suspect_product_active_ingredients, all_of(organ_systems)) %>%
+    pivot_longer(cols = all_of(organ_systems), names_to = "Organ_System", values_to = "Reaction") %>%
+    filter(!is.na(Reaction)) %>%
+    group_by(suspect_product_active_ingredients, Reaction) %>%
+    summarise(Count = n(), .groups = 'drop') %>%
+    mutate(Percent = round((Count / nrow(mydata)) * 100, 1)) %>%
+    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))
+
+# Pivot the summary to have reactions as rows and drugs as columns, combining counts and percentages
+organ_system_pivoted <- organ_system_summary %>%
+    select(suspect_product_active_ingredients, Reaction, Value) %>%
+    pivot_wider(names_from = suspect_product_active_ingredients, values_from = Value, values_fill = list(Value = "0"))
+
+# Create a huxtable table
+ht <- huxtable::as_hux(organ_system_pivoted)
+
+# Add styling to the table
+ht <- ht %>%
+    set_bold(1, 1:ncol(ht)) %>%  # Bold the first row (header)
+    set_caption("Immune Checkpoint Inhibitors by Organ System and Reaction") %>%
+    set_width(1)                  # Set column width to make the table more compact
+
+# Display the table
+ht
+
+#Save the table
+ht_path = here("results", "tables", "summarytable1.rds")
+saveRDS(ht, file = ht_path)
+
 ```
@@
 ```{r}
-p2 <- summary_data %>%
-    ggplot(aes(x = reorder(Category, Count), y = Count)) + # Reorder by Count ascending
+# Define the organ system columns
+organ_systems <- c("organ_system_1", "organ_system_2", "organ_system_3", "organ_system_4", "organ_system_5")
+
+# Create a summary dataset for reactions across all organ systems
+organ_system_summary <- mydata %>%
+    select(all_of(organ_systems)) %>%
+    pivot_longer(cols = everything(), names_to = "Organ_System", values_to = "Reaction") %>%
+    filter(!is.na(Reaction)) %>%
+    group_by(Reaction) %>%
+    summarise(Count = n(), .groups = 'drop') %>%
+    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
+    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))
+
+# View the summary dataset
+organ_system_summary
+
+# Plot the reaction types with counts and percentages
+p2 <- organ_system_summary %>%
+    ggplot(aes(x = reorder(Reaction, Count), y = Count)) +  # Reorder by Count
     geom_col(fill = "#1f77b4") +
@@
         title = "Bar Plot of Reaction Types",
-        x = "Organ System",
+        x = "Reaction Type",
         y = "Count",
@@
     theme_minimal() +
-    geom_text(aes(label = Value), hjust = -0.1, size = 3) + # Adjust label position
-    coord_flip() # Flip the axes
+    geom_text(aes(label = Value), hjust = 1.1, size = 3) +  # Adjust label position to avoid overlap
+    coord_flip() +  # Flip the axes to make it horizontal
+    theme(
+        plot.margin = margin(10, 50, 10, 10)  # Add margin to the right side to avoid label cut-off
+    )
 

### 2025-03-14 code/processing-code/processingfile-v1.qmd
Part 3 of MADA Project

@@
 
+#Reactions with irAES
 ici_data_subset <- ici_data1 %>%
   filter(str_detect(tolower(reactions), paste(tolower(reactions_of_interest), collapse = "|")))
+
+#Reactions with irAEs
+ici_data_non_irae <- ici_data1 %>%
+  filter(!str_detect(tolower(reactions), paste(tolower(reactions_of_interest), collapse = "|")))
 
@@
 
-We have 368 reactions.
+We have 369 reactions.
 The next step is to characterize and categorize the irAEs.
@@
 
+The next step is to remove any duplicates. We will remove reports that have identical age, weight, FDA received date, country of occurrence, suspected active ingredient, and reported reaction.
+
+```{r}
+# Find duplicates based on specified columns
+duplicates <- ici_data_cleaned %>%
+  group_by(suspect_product_active_ingredients, sex, patient_age, patient_weight, reactions, initial_fda_received_date, country_where_event_occurred) %>%
+  filter(n() > 1) %>%
+  ungroup()
+
+# Display the duplicates
+print(duplicates)
+```
+
+We see that there are 4 duplicates of 1 case.
+We remove these.
+
+```{r}
+ici_data_cleaned = ici_data_cleaned %>%
+  distinct(suspect_product_active_ingredients, sex, patient_age, patient_weight, reactions, initial_fda_received_date, country_where_event_occurred, .keep_all = TRUE)
+```
+
 Now, the next step is to remove columns that are not required.
@@
 ici_data_final <- ici_data_final %>%
-  mutate(country_where_event_occurred = ifelse(country_where_event_occurred != "US", "Outside US", country_where_event_occurred))
+  mutate(country_where_event_occurred = case_when(
+    country_where_event_occurred == "US" ~ "US",
+    country_where_event_occurred != "US" & country_where_event_occurred != "Not Specified" ~ "Outside US",
+    TRUE ~ country_where_event_occurred
+  ))
 
 ```
+
+The next step is to classify the irAEs by the System Organ Class (SOC) according to MedDRA.
 
@@
 # Define lists of reaction keywords for each organ system
-dermatological_reactions <- c("rash", "pruritus", "dermatitis", "psoriasis")
-gastrointestinal_reactions <- c("colitis", "hepatitis", "pancreatitis", "enterocolitis", "ileitis", "ileocolitis")
-neurological_reactions <- c("myasthenia", "graves", "myelitis", "encephalitis", "neuritis", "peripheral neuropathy")
-endocrinological_reactions <- c("hyperthyroidism", "hypothyroidism", "thyroiditis", "thyrotoxicosis", "adrenitis", "adrenal", "hypogonadism", "hyperparathyroidism", "hypercortisolism", "hypophysitis", "hypoparathyroidism", "hypocortisolism", "autoimmune diabetes")
-cardiovascular_reactions <- c("myocarditis", "pericarditis")
+Skin_and_subcutaneous_tissue_disorders <- c("rash", "pruritus", "dermatitis", "psoriasis")
+Gastrointestinal_disorders <- c("colitis", "hepatitis", "pancreatitis", "enterocolitis", "ileitis", "ileocolitis")
+Nervous_system_disorders <- c("myasthenia", "graves", "myelitis", "encephalitis", "neuritis", "peripheral neuropathy")
+Endocrine_disorders <- c("hyperthyroidism", "hypothyroidism", "thyroiditis", "thyrotoxicosis", "adrenitis", "adrenal", "hypogonadism", "hyperparathyroidism", "hypercortisolism", "hypophysitis", "hypoparathyroidism", "hypocortisolism", "autoimmune diabetes")
+Cardiac_disorders <- c("myocarditis", "pericarditis", "myocarditis")
+Respiratory_thoracic_and_mediastinal_disorders <- c("pneumonitis")
+Renal_and_urinary_disorders <- c("pyelonephritis", "nephritis")
 
-# Assuming your data is stored in 'ici_data_final' and reactions are in reaction_1 to reaction_5
+# Function to classify reactions into organ systems
+classify_reaction <- function(reaction) {
+  case_when(
+    !is.na(reaction) & grepl(paste(Skin_and_subcutaneous_tissue_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Skin",
+    !is.na(reaction) & grepl(paste(Gastrointestinal_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Gastrointestinal",
+    !is.na(reaction) & grepl(paste(Nervous_system_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Neurological",
+    !is.na(reaction) & grepl(paste(Endocrine_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Endocrinological",
+    !is.na(reaction) & grepl(paste(Cardiac_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Cardiovascular",
+    !is.na(reaction) & grepl(paste(Respiratory_thoracic_and_mediastinal_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Respiratory",
+    !is.na(reaction) & grepl(paste(Renal_and_urinary_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Renal",
+    !is.na(reaction) ~ "Other",  # If not NA and no match, classify as Other
+    TRUE ~ NA_character_  # If NA, keep as NA
+  )
+}
+
+# Apply the function to each reaction column
 ici_data_final <- ici_data_final %>%
   mutate(
-    organ_system_1 = case_when(
-      !is.na(reaction_1) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Skin",
-      !is.na(reaction_1) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Gastrointestinal",
-      !is.na(reaction_1) & grepl(paste(neurological_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Neurological",
-      !is.na(reaction_1) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Endocrinological",
-      !is.na(reaction_1) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Cardiovascular",
-      !is.na(reaction_1) ~ "Other",  # If not NA and no match, classify as Other
-      TRUE ~ NA_character_  # If NA, keep as NA
-    ),
-
-    organ_system_2 = case_when(
-      !is.na(reaction_2) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Skin",
-      !is.na(reaction_2) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Gastrointestinal",
-      !is.na(reaction_2) & grepl(paste(neurological_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Neurological",
-      !is.na(reaction_2) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Endocrinological",
-      !is.na(reaction_2) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Cardiovascular",
-      !is.na(reaction_2) ~ "Other",
-      TRUE ~ NA_character_
-    ),
-
-    organ_system_3 = case_when(
-      !is.na(reaction_3) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Skin",
-      !is.na(reaction_3) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Gastrointestinal",
-      !is.na(reaction_3) & grepl(paste(neurological_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Neurological",
-      !is.na(reaction_3) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Endocrinological",
-      !is.na(reaction_3) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Cardiovascular",
-      !is.na(reaction_3) ~ "Other",
-      TRUE ~ NA_character_
-    ),
-
-    organ_system_4 = case_when(
-      !is.na(reaction_4) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Skin",
-      !is.na(reaction_4) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Gastrointestinal",
-      !is.na(reaction_4) & grepl(paste(neurological_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Neurological",
-      !is.na(reaction_4) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Endocrinological",
-      !is.na(reaction_4) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Cardiovascular",
-      !is.na(reaction_4) ~ "Other",
-      TRUE ~ NA_character_
-    ),
-
-    organ_system_5 = case_when(
-      !is.na(reaction_5) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Skin",
-      !is.na(reaction_5) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Gastrointestinal",
-      !is.na(reaction_5) & grepl(paste(neurological_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Neurological",
-      !is.na(reaction_5) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Endocrinological",
-      !is.na(reaction_5) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Cardiovascular",
-      !is.na(reaction_5) ~ "Other",
-      TRUE ~ NA_character_
-    )
+    organ_system_1 = classify_reaction(reaction_1),
+    organ_system_2 = classify_reaction(reaction_2),
+    organ_system_3 = classify_reaction(reaction_3),
+    organ_system_4 = classify_reaction(reaction_4),
+    organ_system_5 = classify_reaction(reaction_5)
   )
@@
 head(ici_data_final)
+
+```
+Now, let us split the outcomes column.
+
+```{r}
+library(tidyr)
+
+ici_data_final <- ici_data_final %>%
+  separate(outcomes, into = paste0("outcome_", 1:2), sep = ";", remove = FALSE, extra = "drop")
+
+# If outcome is died, removing other outcomes for that report
+ici_data_final$outcome_1[!is.na(ici_data_final$outcome_1) & ici_data_final$outcome_2 == "Died"] <- NA
+ici_data_final$outcome_2[!is.na(ici_data_final$outcome_2) & ici_data_final$outcome_1 == "Died"] <- NA
+
+#If outcome is hospitalized, removing other outcomes for that report
+ici_data_final$outcome_1[!is.na(ici_data_final$outcome_1) & ici_data_final$outcome_2 == "Hospitalized"] <- NA
+ici_data_final$outcome_2[!is.na(ici_data_final$outcome_2) & ici_data_final$outcome_1 == "Hospitalized"] <- NA
+
+#If outcome is life threatening, removing other outcomes for that report
+ici_data_final$outcome_1[!is.na(ici_data_final$outcome_1) & ici_data_final$outcome_2 == "Life Threatening"] <- NA
+ici_data_final$outcome_2[!is.na(ici_data_final$outcome_2) & ici_data_final$outcome_1 == "Life Threatening"] <- NA
+
+#If outcome is disabled, removing other outcomes for that report
+ici_data_final$outcome_1[!is.na(ici_data_final$outcome_1) & ici_data_final$outcome_2 == "Disabled"] <- NA
+ici_data_final$outcome_2[!is.na(ici_data_final$outcome_2) & ici_data_final$outcome_1 == "Disabled"] <- NA
+
+#Merge the columns
+ici_data_final$outcomes <- coalesce(ici_data_final$outcome_1, ici_data_final$outcome_2)  # Merge outcome_1 and outcome_2
+ici_data_final$outcome_1 <- NULL  # Remove outcome_1 column
+ici_data_final$outcome_2 <- NULL  # Remove outcome_2 column
+
+# View the result
+View(ici_data_final)
 
@@
 ```
+
+Since we will need the number of reactions other than irAEs for the further analysis, we summarize that information here.
+
+```{r}
+glimpse(ici_data_non_irae)
+```
+
+First, let us split the reactions column
+```{r}
+library(tidyr)
+
+ici_data_non_irae <- ici_data_non_irae %>%
+  separate(reactions, into = paste0("reaction_", 1:15), sep = ";", remove = FALSE, extra = "drop")
+
+# View the result
+View(ici_data_non_irae)
+
+reaction_columns <- paste0("reaction_", 1:15)
+```
+
+We remove any duplicates.
+```{r}
+duplicates1 <- ici_data_non_irae %>%
+  group_by(suspect_product_active_ingredients, sex, patient_age, patient_weight, reactions, initial_fda_received_date, country_where_event_occurred) %>%
+  filter(n() > 1) %>%
+  ungroup()
+
+# Display the duplicates
+print(duplicates1)
+```
+
+We see multiple duplicates. Let us remove them.
+```{r}
+ici_data_clean_non_irae = ici_data_non_irae %>%
+  distinct(suspect_product_active_ingredients, sex, patient_age, patient_weight, reactions, initial_fda_received_date, country_where_event_occurred, .keep_all = TRUE)
+
+View(ici_data_clean_non_irae)
+```
+
+We will need this data in the disproportionality analysis.
+
+We save this data as well.
+
+```{r}
+processeddata1 = ici_data_clean_non_irae
+save_data_location <- here::here("data","processed-data","processeddata1.rds")
+saveRDS(processeddata1, file = save_data_location)
+```

### 2025-03-14 products/manuscript/Manuscript.qmd
Fixed the rendering error
Part 3 of MADA Project

@@
 
-*Write a summary of your project.* *This part has not been completed. The source code has been left as is*
+*Background: Bladder cancer is the ninth most common type of cancer worldwide.
+Treatment options include surgery, chemotherapy and immunotherapy.
+Immune checkpoint inhibitors (ICIs) are a class of immunotherapeutic agents recommended for the treatment of bladder cancer.
+ICIs have been associated with certain immune-related adverse events (irAEs)*
+
+*Methods: Reports in the FAERS dashbaord through December 2024 were analyzed for irAEs to four ICIs recommended in bladder cancer - atezolizumab, avelumab, nivolumab, and pembrolizumab.
+Standardized queries were used to identify cases of rash, pruritus, hypothyroidism, hyperthyroidism, colitis, hepatitis, and nephritis.
+Descriptive and disproportionality analyses were performed to detect safety signals.*
+
+*Results: To be added in Part 4*
 
@@
 Following this, the search was limited to cases of bladder cancer.
-A total of 1,886 reactions were obtained.*
+A total of 2948 reactions were obtained.*
 
@@
 *Reactions with concomitant suspect products, combination products, and those which were not irAEs were removed.
-This resulted in a total of 369 cases.
+This resulted in a total of 365 cases.
 The details are available in the processing code file.*
+
+*The processing was done according to @fig-processing1.*
+
+```{r}
+#| label: fig-processing1
+#| fig-cap: "Study Flowchart"
+#| echo: FALSE
+knitr::include_graphics(here("assets","Figure1.png"))
+```
 
@@
 
-*The second part of the analysis will involve a disproportionality analysis. This analysis uses the following formula to calculate the reported odds ratio (ROR).*
-
-*a = irAES reported with ICIs* *b = irAEs reported with all other drugs* *c = All other AEs reported with ICIs* *d = All other AEs reported with other drugs*
-
-*ROR = a/c / b/d*
-
-*Other details in the analysis will be finalized later.*
+*The second part of the analysis will involve a disproportionality analysis.
+This analysis employs Proportional Reporting Ratio (PRR), Reporting Odds Ratio (ROR), and Information Component (IC).
+Further details will be added in Part 4..*
 
@@
 # Results
-
-*This part has not been completed. The source code has been left as is*
 
@@
 knitr::kable(resulttable)
+
+```
+
+```{r}
+
 ```
@@
 ## Basic statistical analysis
+
+*@tbl-summarytable1 shows the number of reactions by organ system and the ICI.*
+
+```{r}
+#| label: tbl-summarytable1
+#| tbl-cap: "Table 2: Reactions by organ system and drug"
+#| echo: FALSE
+resulttable1=readRDS("../../results/tables/summarytable1.rds")
+knitr::kable(resulttable1)
+```
 
@@
 
-*This part has not yet been edited.*
-
-*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here.*
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
+*The following tables show the results from the disproportionality analysis.*
 
 ```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
+#| label: tbl-atezolizumab
+#| tbl-cap: "Table 1: Atezolizumab Reactions by Organ System"
 #| echo: FALSE
-resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
-knitr::kable(resulttable2)
+table_atezolizumab <- readRDS("../../results/tables/table_atezolizumab.rds")
+knitr::kable(table_atezolizumab)
 ```
+
+```{r}
+#| label: tbl-avelumab
+#| tbl-cap: "Table 2: Avelumab Reactions by Organ System"
+#| echo: FALSE
+table_avelumab <- readRDS("../../results/tables/table_avelumab.rds")
+knitr::kable(table_avelumab)
+```
+
+```{r}
+#| label: tbl-nivolumab
+#| tbl-cap: "Table 3: Nivolumab Reactions by Organ System"
+#| echo: FALSE
+table_nivolumab <- readRDS("../../results/tables/table_nivolumab.rds")
+knitr::kable(table_nivolumab)
+```
+
+```{r}
+#| label: tbl-pembrolizumab
+#| tbl-cap: "Table 4: Pembrolizumab Reactions by Organ System"
+#| echo: FALSE
+table_pembrolizumab <- readRDS("../../results/tables/table_pembrolizumab.rds")
+knitr::kable(table_pembrolizumab)
+
+```
+
+*The PRR and ROR when all irAEs are considered is greater than 1 for pembrolizumab (PRR: 1.8 (1.52, 2.14), ROR: 2.07 (1.69, 2.55)).
+Similarly, the IC is greater than 0 (0.36 (0.29, 0.44)).
+Thus, there is a safety signal associated with pembrolizumab for all irAEs.*
+
+*When considering individual irAEs, pembrolizumab also has a safety signal for rash (PRR: 2.11 (1.17, 3.8), ROR: 2.14 (1.18, 3.9), IC: 0.44 (0.17, 0.71).*
+
+*Atezolizumab has a safety signal for hyperthyroidism (PRR: 3.51 (1.07, 11.45), ROR: 3.53 (1.07, 11.61), IC:	1.1 (0.33, 1.87).
+There also may be a signal for nephritis (PRR: 3.65 (0.98, 13.56), ROR: 3.68 (0.98, 13.73), IC:	1.12 (0.29, 1.96), though the 95% CI for PRR and ROR is not greater than 1 (though very close to 1).*
+
+*Similarly, avelumab may has a safety signal for hypothyroidism (PRR: 3.05 (0.98, 9.51), ROR:	3.08 (0.97, 9.75), IC:	1.32 (0.12, 2.52)).*
+
+*No significant safety signals were found for nivolumab.*
 

### 2025-04-04 code/analysis-code/statistical-analysis.qmd
Part 4 of MADA Project

@@
 
-# Pivot data to get reactions as rows and drugs as grouped columns with subcolumns for PRR, ROR, and IC
+# Pivot data to get reactions as rows and drugs as grouped columns
 summary_table <- all_results %>%
-  pivot_wider(names_from = Drug, values_from = c(PRR, ROR, IC),
-              names_glue = "{Drug} {.value} (95% CI)")
-
-# Check the structure of the table and columns
-colnames(summary_table)
-
-# Highlight significant values using conditional bolding (PRR > 1, ROR > 1, IC > 0)
-summary_table <- summary_table %>%
-  mutate(across(starts_with("PRR"), ~ ifelse(. > 1 & grepl("\\(1", .), "bold", .))) %>%
-  mutate(across(starts_with("ROR"), ~ ifelse(. > 1 & grepl("\\(1", .), "bold", .))) %>%
-  mutate(across(starts_with("IC"), ~ ifelse(. > 0 & grepl("\\(0", .), "bold", .)))
-
-# Split the summary_table into 4 separate tables for each drug
-# Table for Atezolizumab
-table_atezolizumab <- summary_table %>%
-  select(Reaction, starts_with("Atezolizumab")) %>%
-  as_hux() %>%
-  set_bold(1, 1:ncol(.)) %>%
-  set_caption("Atezolizumab Drug-Reaction Analysis") %>%
-  set_width(1)
-
-# Apply bolding for significant values for Atezolizumab
-for (i in 2:ncol(table_atezolizumab)) {
-  table_atezolizumab <- table_atezolizumab %>%
-    set_bold(which(table_atezolizumab[, i] == "bold"), i)
-}
-
-# Table for Avelumab
-table_avelumab <- summary_table %>%
-  select(Reaction, starts_with("Avelumab")) %>%
-  as_hux() %>%
-  set_bold(1, 1:ncol(.)) %>%
-  set_caption("Avelumab Drug-Reaction Analysis") %>%
-  set_width(1)
-
-# Apply bolding for significant values for Avelumab
-for (i in 2:ncol(table_avelumab)) {
-  table_avelumab <- table_avelumab %>%
-    set_bold(which(table_avelumab[, i] == "bold"), i)
-}
-
-# Table for Nivolumab
-table_nivolumab <- summary_table %>%
-  select(Reaction, starts_with("Nivolumab")) %>%
-  as_hux() %>%
-  set_bold(1, 1:ncol(.)) %>%
-  set_caption("Nivolumab Drug-Reaction Analysis") %>%
-  set_width(1)
-
-# Apply bolding for significant values for Nivolumab
-for (i in 2:ncol(table_nivolumab)) {
-  table_nivolumab <- table_nivolumab %>%
-    set_bold(which(table_nivolumab[, i] == "bold"), i)
-}
-
-# Table for Pembrolizumab
-table_pembrolizumab <- summary_table %>%
-  select(Reaction, starts_with("Pembrolizumab")) %>%
-  as_hux() %>%
-  set_bold(1, 1:ncol(.)) %>%
-  set_caption("Pembrolizumab Drug-Reaction Analysis") %>%
-  set_width(1)
-
-# Apply bolding for significant values for Pembrolizumab
-for (i in 2:ncol(table_pembrolizumab)) {
-  table_pembrolizumab <- table_pembrolizumab %>%
-    set_bold(which(table_pembrolizumab[, i] == "bold"), i)
-}
-
-# Display each table
+  pivot_wider(
+    names_from = Drug, 
+    values_from = c(PRR, ROR, IC),
+    names_glue = "{Drug} {.value} (95% CI)"
+  )
+
+# Function to create formatted tables with WORKING conditional formatting
+create_drug_table <- function(drug_name, summary_data) {
+  # Select relevant columns
+  drug_table <- summary_data %>%
+    select(Reaction, starts_with(drug_name)) %>%
+    as_hux(add_colnames = TRUE)
+
+  # Basic formatting
+  drug_table <- drug_table %>%
+    set_bold(1, everywhere, TRUE) %>%          # Bold headers
+    set_align(everywhere, everywhere, "center") %>%
+    set_all_borders(brdr(1, "solid")) %>%      # Solid borders
+    set_all_padding(4) %>%                     # Cell padding
+    set_font_size(10) %>%                      # Font size
+    set_width(0.9)                             # Table width
+
+  # Add caption as merged header row
+  drug_table <- insert_row(drug_table, 
+                         c(paste(drug_name, "Drug-Reaction Analysis"), 
+                           rep("", ncol(drug_table)-1)), 
+                         after = 0) %>%
+    set_colspan(1, 1, ncol(drug_table)) %>%
+    set_bold(1, everywhere, TRUE) %>%          # Bold title
+    set_background_color(1, everywhere, "gray95") %>%
+    set_font_size(1, everywhere, 12)
+
+  # Apply conditional formatting - SIMPLIFIED AND WORKING VERSION
+  if (ncol(drug_table) > 1) {
+    for (i in 2:ncol(drug_table)) {
+      # Get the column type (PRR, ROR, or IC)
+      col_type <- colnames(drug_table)[i]
+
+      # Process each row (skip header row)
+      for (j in 2:nrow(drug_table)) {
+        cell_value <- as.character(drug_table[j, i])
+
+        # Extract the first number (point estimate)
+        point_est <- as.numeric(gsub("([0-9.]+).*", "\\1", cell_value))
+
+        # Extract the first number in parentheses (CI lower bound)
+        ci_lower <- as.numeric(gsub(".*\\(([0-9.]+).*", "\\1", cell_value))
+
+        # Check significance
+        if (!is.na(point_est) && !is.na(ci_lower)) {
+          if (grepl("PRR|ROR", col_type) && point_est > 1 && ci_lower > 1) {
+            drug_table <- set_bold(drug_table, j, i, TRUE)
+          } else if (grepl("IC", col_type) && point_est > 0 && ci_lower > 0) {
+            drug_table <- set_bold(drug_table, j, i, TRUE)
+          }
+        }
+      }
+    }
+  }
+
+  return(drug_table)
+}
+
+# Create all tables
+table_atezolizumab <- create_drug_table("Atezolizumab", summary_table)
+table_avelumab <- create_drug_table("Avelumab", summary_table)
+table_nivolumab <- create_drug_table("Nivolumab", summary_table)
+table_pembrolizumab <- create_drug_table("Pembrolizumab", summary_table)
+
+# Display tables
 table_atezolizumab
@@
 table_pembrolizumab 
-
 ```
@@
 ```
+
+Now, we use machine learning methods (random forest) to assess severity of reaction based on age, sex and drug.
+
+```{r}
+library(tidyverse)
+library(randomForest) # For Random Forest
+library(xgboost)     # For XGBoost
+library(caret)       # For preprocessing
+
+glimpse(mydata)
+
+# Check factor levels and missingness
+summary(mydata[c("suspect_product_active_ingredients", "sex", "patient_age", "serious")])
+
+# Class distribution
+table(mydata$serious)  # Check for imbalance (e.g., 70% non-serious vs. 30% serious)
+```
+
+We see missing data and imbalance of outcomes. First, we impute age for the missing data.
+```{r}
+library(tidyverse)
+library(randomForest)
+
+df_clean <- mydata %>%
+  mutate(
+    # Convert to factors (critical for ML)
+    suspect_product_active_ingredients = factor(suspect_product_active_ingredients),
+    sex = factor(sex),
+    serious = factor(serious),
+
+    # Impute missing age with median (stratified by sex to preserve trends)
+    patient_age = ifelse(
+      is.na(patient_age),
+      median(patient_age, na.rm = TRUE),
+      patient_age
+    )
+  )
+```
+
+Now, we split the data:
+```{r}
+library(caret)
+set.seed(123)
+
+# Stratified split (70% train, 30% test)
+train_index <- createDataPartition(
+  df_clean$serious, 
+  p = 0.7, 
+  list = FALSE,
+  times = 1
+)
+
+train_data <- df_clean[train_index, ]
+test_data <- df_clean[-train_index, ]
+
+# Verify class balance in both sets
+table(train_data$serious)  # Should be ~70% of original counts
+table(test_data$serious)   # Should be ~30%
+```
+We can see the split has happened correctly, and the imbalance is there. Let's use inverse weighting to deal with the imbalance, and then fit a random forest model.
+
+```{r}
+library(randomForest)
+
+# 1. Verify factor levels (critical!)
+levels(train_data$serious)  # Must return: [1] "Non-Serious" "Serious"
+
+# 2. Calculate class weights (EXACT names and case)
+class_weights <- c(
+  "Non-Serious" = sum(train_data$serious == "Serious") / nrow(train_data),  # Majority class weight
+  "Serious" = sum(train_data$serious == "Non-Serious") / nrow(train_data)    # Minority class weight
+)
+```
+
+Now, let's build a random forest model using cross-validation.
+
+```{r}
+library(randomForest)
+library(caret)
+library(pROC)
+
+# Ensure `serious` is a factor
+train_data$serious <- as.factor(train_data$serious)
+
+# Define class weights based on inverse frequency
+class_counts <- table(train_data$serious)
+class_weights <- 1 / class_counts  # Inverse class frequency weighting
+class_weights <- setNames(as.numeric(class_weights), names(class_counts))  # Ensure correct names
+
+# Perform 5-fold cross-validation
+set.seed(123)
+k_folds <- 5
+folds <- createFolds(train_data$serious, k = k_folds, list = TRUE)
+
+cv_results <- lapply(folds, function(fold_idx) {
+  train_fold <- train_data[-fold_idx, ]
+  test_fold  <- train_data[fold_idx, ]
+
+  # Train Random Forest model
+  rf_model <- randomForest(
+    serious ~ suspect_product_active_ingredients + sex + patient_age,
+    data = train_fold,
+    classwt = class_weights,  # Correct class weights
+    strata = train_fold$serious,
+    sampsize = rep(min(table(train_fold$serious)), 2),
+    ntree = 500,
+    importance = TRUE,
+    na.action = na.roughfix
+  )
+
+  # Make predictions
+  preds <- predict(rf_model, newdata = test_fold, type = "response")
+  probs <- predict(rf_model, newdata = test_fold, type = "prob")[, 2]  # Get probabilities for ROC
+
+  # Compute confusion matrix
+  cm <- confusionMatrix(preds, test_fold$serious)
+
+  # Compute evaluation metrics
+  accuracy  <- cm$overall["Accuracy"]
+  precision <- cm$byClass["Precision"]
+  recall    <- cm$byClass["Recall"]
+  f1_score  <- 2 * (precision * recall) / (precision + recall)
+
+  # Compute AUC-ROC
+  roc_curve <- roc(as.numeric(test_fold$serious), probs)
+  auc_value <- auc(roc_curve)
+
+  return(c(accuracy, precision, recall, f1_score, auc_value))
+})
+
+library(huxtable)
+
+# Assuming cv_results_df is already created as shown in your code
+cv_results_df <- as.data.frame(do.call(rbind, cv_results))
+colnames(cv_results_df) <- c("Accuracy", "Precision", "Recall", "F1-Score", "AUC-ROC")
+
+# Compute average metrics and create a summary table
+avg_metrics <- colMeans(cv_results_df)
+metrics_table <- data.frame(
+  Metric = names(avg_metrics),
+  Value = round(avg_metrics, 4)  # Round to 4 decimal places
+)
+
+library(huxtable)
+
+# Create the metrics data frame
+metrics_table <- data.frame(
+  Metric = c("Accuracy", "Precision", "Recall", "F1-Score", "AUC-ROC"),
+  Value = round(avg_metrics, 4)  # Using your computed avg_metrics
+)
+
+clean_table <- as_hux(metrics_table) %>%
+  # Formatting
+  set_bold(1, everywhere) %>%           # Bold header row only
+  set_align(everywhere, everywhere, "center") %>%  # Center all cells
+  set_width(0.6) %>%                    # Set table width
+  set_font_size(10) %>%                 # Font size
+  # Borders (the key fix)
+  set_all_borders(1) %>%                # All cell borders (1 = thin line)
+  set_outer_borders(2)                  # Thicker outer border (2 = thicker)
+
+# Print the table
+clean_table
+
+```
+```{r}
+# Define paths for saving each table
+clean_table_path <- here("results", "tables", "randomforesttable.rds")
+
+# Save all four tables to their respective paths
+saveRDS(clean_table, file = clean_table_path)
+```
+
+We can see that the accuracy is quite low (25.3%) - thus, the model is only able to make correct predictions 25% of the time. 
+There are many false positives (Precision: 16.61%).
+The model is able to capture 87.78% of positives, but at the cost of correctly predicting negatives.
+The AUC-ROC is low, thus the model is not able to distinguish between positives and negatives effectively.
+
+If we had a larger dataset, or more variables, we might be able to build a better model.
+I will explore more modifications and complex methods in part 5.

### 2025-04-04 products/manuscript/Manuscript.qmd
Part 4 of MADA Project

--- 2025-02-21 products/manuscript/Manuscript.qmd
@@
 library(here)
+library(huxtable)
 library(knitr)
@@
 
-*Write a summary of your project.* *This part has not been completed. The source code has been left as is*
+Background: Bladder cancer is the ninth most common type of cancer worldwide.
+Treatment options include surgery, chemotherapy and immunotherapy.
+Immune checkpoint inhibitors (ICIs) are a class of immunotherapeutic agents recommended for the treatment of bladder cancer.
+ICIs have been associated with certain immune-related adverse events (irAEs)
+
+Methods: Reports in the FAERS dashbaord through December 2024 were analyzed for irAEs to four ICIs recommended in bladder cancer - atezolizumab, avelumab, nivolumab, and pembrolizumab.
+Standardized queries were used to identify cases of rash, pruritus, hypothyroidism, hyperthyroidism, colitis, hepatitis, and nephritis.
+Descriptive and disproportionality analyses were performed to detect safety signals.
+
+Results: Out of 2948 cases identified from FAERS, 365 cases with irAEs were analyzed.
+The largest number of cases were reported for pembrolizumab (n=307), followed by atezolizumab (n = 91), avelumab (n = 41), and nivolumab (n =37).
+Pembrolizumab was found to have safety signals for all irAEs considered together (PRR: 1.8 (1.52, 2.14), ROR: 2.07 (1.69, 2.55), IC:	0.36 (0.29, 0.44)).
+A safety signal was also found for rash (PRR: 2.11 (1.17, 3.8), ROR:	2.14 (1.18, 3.9), IC:	0.44 (0.17, 0.71)).
+Safety signals for atezolizumab included hyperthyroidism and nephritis, while hypothyroidism was identified as a concern for avelumab.
+
+Conclusion: Immune-related adverse events are found to be associated with immune checkpoint inhibitor therapy in bladder cancer.
+Pembrolizumab appears to have the highest risk, though this may be due to its early approval (2014).
+Appropriate treatment guidelines must be developed to address the management of irAEs associated with ICI therapy in bladder cancer.
 
@@
 
-*Bladder cancer is the ninth most common type of cancer, according to a 2022 report published by the International Agency for Research on Cancer [@Bray2024].
-The American Cancer Society estimates approximately 85,000 new bladder cancer cases, and approximately 17,000 bladder cancer deaths in the United States in 2025.*
+Bladder cancer is the ninth most common type of cancer, according to a 2022 report published by the International Agency for Research on Cancer [@Bray2024].
+The American Cancer Society estimates approximately 85,000 new bladder cancer cases, and approximately 17,000 bladder cancer deaths in the United States in 2025.
 
-*The cancer is more common in males than in females, with a male:female ratio of about 4:1[@Leslie2025-lp].
+The cancer is more common in males than in females, with a male-female ratio of about 4:1[@Leslie2025-lp].
 Risk factors for bladder cancer include smoking and occupational chemical exposures (in particular, aromatic amines and polycyclic aromatic hydrocarbons).
-Certain dietary factors, impaired microbiome, and pelvic radiotherapy may also increase the risk of bladder cancer [@Jubber2023].*
+Certain dietary factors, impaired microbiome, and pelvic radiotherapy may also increase the risk of bladder cancer [@Jubber2023].
 
-*Treatment of bladder cancer is multifaceted and depends on the type of tumor.
-Surgical resection, adjuvant and intravesical chemotherapy, Bacillus Calmette–Guérin (BCG) immunotherapy, and systemic chemotherapy are the different treatment options available [@Leslie2025-lp].*
+Treatment of bladder cancer is multifaceted and depends on the type of tumor.
+Surgical resection, adjuvant and intravesical chemotherapy, Bacillus Calmette–Guérin (BCG) immunotherapy, and systemic chemotherapy are the different treatment options available [@Leslie2025-lp].
 
-*According to the National comprehensive Cancer Network (NCCN) guidelines, immunotherapy with immune checkpoint inhibitors (ICIs) is recommended at various stages of bladder cancer treatment.
-ICIs such as pembrolizumab, atezolizumab, nivolumab, and avelumab are FDA approved and recommended for the adjuvant and first and second-line treatment of bladder cancer [@Flaig2024].*
+According to the National Comprehensive Cancer Network (NCCN) guidelines, immunotherapy with immune checkpoint inhibitors (ICIs) is recommended at various stages of bladder cancer treatment.
+ICIs such as pembrolizumab, atezolizumab, nivolumab, and avelumab are FDA approved and recommended for the adjuvant and first and second-line treatment of bladder cancer [@Flaig2024].
 
-*ICIs are associated with various adverse effects, particularly with immune-related adverse events (irAEs) such as pnemonitis, hepatitis, colitis, hypothyroidism, dermatitis, and others [@Yin2023].This study aims to analyze real-world adverse event data from the FDA Adverse Event Reporting System associated with ICI therapy in bladder cancer.*
+ICIs are associated with various adverse effects, particularly with immune-related adverse events (irAEs) such as pneumonitis, hepatitis, colitis, hypothyroidism, dermatitis, and others [@Yin2023].This study aims to analyze real-world adverse event data from the FDA Adverse Event Reporting System associated with ICI therapy in bladder cancer.
 
@@
 
-*The United States Food and Drug Administration (FDA) maintains a dashboard of adverse event reports in the FDA Adverse Event Reporting System (FAERS) [@us2018fda].
-FAERS is a publicly available dashboard of safety reports submitted by patients, healthcare professionals, and pharmaceutical companies.*
+The United States Food and Drug Administration (FDA) maintains a dashboard of adverse event reports in the FDA Adverse Event Reporting System (FAERS) [@us2018fda].
+FAERS is a publicly available dashboard of safety reports submitted by patients, healthcare professionals, and pharmaceutical companies.
 
-*Since FAERS consists of self-reported data from multiple sources, there may be duplicates, missing data as well as various errors.
-The findings of this study should be considered in context of this data limitation.*
+Since FAERS consists of self-reported data from multiple sources, there may be duplicates, missing data as well as various errors.
+The findings of this study should be considered in context of this data limitation.
 
@@
 
-*The study aims to analyze whether irAEs associated with ICIs generate a signal of disproportionate reporting (SDR).
-The goal is to determine whether irAEs associated with ICIs in baldder cancer are greater than expected, and thus further research and strategies to deal with these adverse events is required.*
+The study aims to analyze whether irAEs associated with ICIs generate a signal of disproportionate reporting (SDR).
+The goal is to determine whether irAEs associated with ICIs in baldder cancer are greater than expected, and thus further research and strategies to deal with these adverse events is required.
 
@@
 
-*Describe your methods.
-That should describe the data, the cleaning processes, and the analysis approaches.
-You might want to provide a shorter description here and all the details in the supplement.* *This part has not been completed.
-The source code has been left as is* *Details have been provided in the data acquisition and statistical analysis parts.*
+The FAERS dashboard was searched by product to obtain cases related to the use of ICIs in bladder cancer.
+This data was then filtered to retain only irAE cases.
+Other adverse events and duplicate cases were removed.
+Following this, the demographic and drug characteristics were summarized.
+
+To assess whether an ICI has a safety signal associated with a particular irAE, three metrics were used:
+1. Proportional reporting ratio (PRR):
+This approach compares the proportion of reports for a specific adverse event with a drug of interest to the proportion of the same event reported for all other drugs.
+
+2. Reporting odds ratio (ROR):
+This metric is obtained by comparing the ratio of a drug's adverse event reports to all other reports for that drug against the same ratio across all drugs.
+
+3. Information component (IC):
+For this metric, the observed vs. expected reporting of a drug-adverse event combination is compared.
+
+A random forest model was fit to distinguish between serious and non-serious cases on the basis of age, sex, and ICI.
 
@@
 
-*In order to obtain the FAERS data, the ICIs were used to search by product.
+In order to obtain the FAERS data, the ICIs were used to search by product.
 Pembrolizumab, atezolizumab, avelumab, and nivolumab were the keywords used.
@@
 Following this, the search was limited to cases of bladder cancer.
-A total of 1,886 reactions were obtained.*
+A total of 2948 reactions were obtained.
 
@@
 
-*The file has been downloaded as an Excel file.
-The code for the analysis would involve loading the file, removing duplicates, assessing missing data, and deciding whether to remove it.* *This will be followed by filtering the data to retain only irAEs.
-While this step could be performed within the FAERS dashboard, this is not always a straightforward process.
-FAERS writes all the reactions associated with an agent in a single column, such that one column may have over 10-15 reactions.
-As such, this step will be performed here.*
+The file was downloaded as an Excel file.
+First, the column were renamed so as to make them easier to use in code.
+Then, products with concomitant drugs were removed, so as to focus only on cases associated with ICIs.
+Then, cases which also had other drugs as the suspect active ingredients were removed.
+This was followed by retaining only the irAEs.
+A list of common irAEs was used.
+This list is available in the supplementary material.
+Any reactions which had the same age, weight, FDA received date, country of occurrence, suspected active ingredient, and reported reaction were considered duplicates and were removed.
+Following this, the reactions were coded according to the MedDRA Preferred Terms (PTs), and classified according to the System Organ Class (SOC).
+The age values were converted to numeric and the countries were re-classified as US, outside US, and not specified.
+The code for the analysis would involve loading the file, removing duplicates, assessing missing data, and deciding whether to remove it. The details are available in the processing code file.
 
-*The FAERS data initially consisted of 2948 reactions associated with ICIs in bladder cancer.*
+The processing was done according to @fig-processing1.
 
-*Reactions with concomitant suspect products, combination products, and those which were not irAEs were removed.
-This resulted in a total of 369 cases.
-The details are available in the processing code file.*
+```{r}
+#| label: fig-processing1
+#| fig-cap: "Study Flowchart"
+#| echo: FALSE
+knitr::include_graphics(here("assets","Figure1.png"))
+```
 
@@
 
-*The first of the analysis will consists of descriptive statistics. Continous variables will be summarized as means, whereas categorical variables will be summarized as counts and percenatges.*
-
-*The second part of the analysis will involve a disproportionality analysis. This analysis uses the following formula to calculate the reported odds ratio (ROR).*
-
-*a = irAES reported with ICIs* *b = irAEs reported with all other drugs* *c = All other AEs reported with ICIs* *d = All other AEs reported with other drugs*
-
-*ROR = a/c / b/d*
-
-*Other details in the analysis will be finalized later.*
+Firstly, descriptive statistics were used to describe the demographic and other characteristics.
+This was followed by creating some descriptive figures.
+The PRR, ROR and IC (along with their 95% CI) were then calculated to assess the drug signal.
+A random forest model was built using age, sex, and ICI, to distinguish between serious and non-serious cases.
 
@@
 
-*This part has not been completed. The source code has been left as is*
-
 ## Exploratory/Descriptive analysis
 
-@tbl-summarytable *provides the patient characteristics, and summary of reactions by type of immune checkpoint inhibitor, and organ system affected.*
+@tbl-summarytable provides the patient characteristics, and summary of reactions by type of immune checkpoint inhibitor, and organ system affected.
 
-*The mean age reported was 66.94±17.65. A large percentage of the reactions occurred in males (70.2%). This may be because bladder cancer is much more common in males than in females.*
+The mean age reported was 66.94±17.65. A large percentage of the reactions occurred in males (70.2%). This may be because bladder cancer is much more common in males than in females.
 
-*Over 60% of the reactions were associated with pembrolizumab. A majority of the reactions were serious (83.7%)*
+Over 60% of the reactions were associated with pembrolizumab. A majority of the reactions were serious (83.7%).
+
+Most of the reaction were reported by healthcare professionals. Since FAERS is a US-based dashboard, many reactions were reported from the US.
 
@@
 #| label: tbl-summarytable
-#| tbl-cap: "Table 1: Descriptive Statistics"
+#| tbl-cap: "Descriptive Statistics"
 #| echo: FALSE
-resulttable=readRDS("../../results/tables/summarytable.rds")
-knitr::kable(resulttable)
+#| warning: FALSE
+#| message: FALSE
+
+library(flextable)
+
+# Load and modify table (all processing hidden)
+demographic_stats <- readRDS(here::here("results", "tables", "summarytable.rds"))[-c(1, 2), ]
+
+# Create and format flextable (visible in output only)
+desc_stats_table <- flextable(demographic_stats) |> 
+  # Borders
+  border_outer(part = "all") |> 
+  border_inner(part = "all") |> 
+  # Formatting
+  bold(part = "header") |> 
+  fontsize(size = 10, part = "all") |> 
+  color(color = "black", part = "all") |> 
+  # Layout
+  set_table_properties(width = 1, layout = "autofit") |>  # Full width
+  align(align = "center", part = "header") |> 
+  padding(padding = 4, part = "all")
+
+# Display table
+desc_stats_table
 ```
@@
 
-*@fig-result shows the number of reactions by drug, along with the seriousness of the reaction.*
+@tbl-summarytable1 shows the number of reactions by organ system and the ICI.
 
 ```{r}
-#| label: fig-result
-#| fig-cap: "Bar chart of outcome of immune checkpoint inhibitors"
+#| label: tbl-summarytable1
+#| tbl-cap: "Reactions by Organ System and Drug"
 #| echo: FALSE
-knitr::include_graphics(here("results","figures","figure1.png"))
+#| warning: FALSE
+#| message: FALSE
+
+library(flextable)
+library(here)
+
+# Load and prepare data
+resulttable1 <- readRDS(here("results", "tables", "summarytable1.rds"))
+resulttable1_df <- as.data.frame(resulttable1)
+colnames(resulttable1_df) <- as.character(resulttable1_df[1,])
+resulttable1_df <- resulttable1_df[-1,]
+
+# Count number of columns
+n_cols <- ncol(resulttable1_df)
+
+# Create and format flextable
+ft <- flextable(resulttable1_df) |> 
+  # Borders
+  border_outer(part = "all") |> 
+  border_inner(part = "all") |> 
+  border_inner_h(part = "all") |> 
+  border_inner_v(part = "all") |> 
+  # Formatting
+  bold(part = "header") |> 
+  fontsize(size = 10, part = "all") |> 
+  color(color = "black", part = "all") |> 
+  # Width control
+  set_table_properties(width = 1, layout = "autofit") |> 
+  # Set relative column widths (adjust proportions as needed)
+  width(width = rep(1, n_cols))  # Equal width for all columns
+
+# Alternative for different column widths:
+# width(width = c(2, rep(1, n_cols-1)))  # First column wider
+
+# Display table
+ft
 ```
-
-*@fig-result1 shows the bar chart by type of organ system involved in the reaction.*
-
-```{r}
-#| label: fig-result1
-#| fig-cap: "Bar chart of types of reactions"
-#| echo: FALSE
-knitr::include_graphics(here("results","figures","figure2.png"))
-```
+Pembrolizumab has the highest number of reactions.
+The largest percentage of reactions are classified as "Other".
+These include reactions which do not belong to the other organ classes, such as myositis.
 
@@
 
-*This part has not yet been edited.*
+For the disproportionality analysis, we need 4 different values:
+a: Number of reports reporting both the drug and the adverse event. 
+b: Number of reports reporting the drug but not the adverse event. 
+c: Number of reports reporting the adverse event but not the drug. 
+d: Number of reports reporting neither the drug nor the adverse event.
 
-*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here.*
+For PRR, the formula used is:
+(a/(a+c)) / (b/(b+d))
 
-Example @tbl-resulttable2 shows a summary of a linear model fit.
+For ROR:
+(a/c)/(b/d)
+
+For IC: 
+log2((a+b+c+d)/(a+c)*(a+b)/n)
+
+The following tables show the results from the disproportionality analysis.
 
 ```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
+#| label: tbl-atezolizumab
+#| tbl-cap: "Atezolizumab Reactions by Organ System"
 #| echo: FALSE
-resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
-knitr::kable(resulttable2)
+#| warning: FALSE
+#| message: FALSE
+
+library(flextable)
+library(officer)
+
+# Load and prepare the table
+table_atezolizumab <- readRDS("../../results/tables/table_atezolizumab.rds") |>
+  (\(x) x[-c(1, 2), ])()  # Remove first two rows using base R
+
+# Create and format flextable
+atezolizumab_table <- flextable(table_atezolizumab) |> 
+  # Basic formatting
+  bold(part = "header") |> 
+  border(part = "all", border = officer::fp_border(color = "black", width = 1)) |> 
+  # Text alignment
+  align(align = "center", part = "header") |> 
+  align(align = "left", part = "body") |> 
+  # Cell padding
+  padding(padding = 4, part = "all") |> 
+  # Font settings
+  fontsize(size = 10, part = "all") |> 
+  # Width control
+  set_table_properties(
+    width = 1,                # 100% of available width
+    layout = "autofit"        # Automatic column adjustment
+  ) |> 
+  # Column width fine-tuning (adjust values as needed)
+  width(j = NULL, width = c(2, rep(1, ncol(table_atezolizumab)-1)))
+
+# Display the table
+atezolizumab_table
 ```
+
+```{r}
+#| label: tbl-avelumab
+#| tbl-cap: "Avelumab Reactions by Organ System"
+#| echo: FALSE
+#| warning: FALSE
+#| message: FALSE
+
+library(flextable)
+library(officer)  # For fp_border()
+
+# Load and prepare data
+avelumab_data <- readRDS("../../results/tables/table_avelumab.rds")[-c(1, 2), ]
+
+# Create and format flextable
+avelumab_table <- flextable(avelumab_data) |> 
+  # Formatting
+  bold(part = "header") |> 
+  # Borders
+  border(part = "all", border = fp_border(color = "black", width = 1)) |> 
+  # Width control
+  set_table_properties(width = 1, layout = "autofit") |> 
+  # Column widths (adjust values as needed)
+  width(width = c(2, rep(1.2, ncol(avelumab_data)-1))) |> 
+  # Additional formatting
+  align(align = "center", part = "header") |> 
+  padding(padding = 4, part = "all") |> 
+  fontsize(size = 10, part = "all")
+
+# Display table 
+avelumab_table
+```
+
+```{r}
+#| label: tbl-nivolumab
+#| tbl-cap: "Nivolumab Reactions by Organ System"
+#| echo: FALSE
+#| warning: FALSE
+#| message: FALSE
+
+library(flextable)
+library(officer)  # Required for fp_border()
+
+# Load and prepare data
+nivolumab_data <- readRDS("../../results/tables/table_nivolumab.rds")[-c(1, 2), ]
+
+# Create and format flextable
+nivolumab_table <- flextable(nivolumab_data) |> 
+  # Basic formatting
+  bold(part = "header") |> 
+  # Complete border styling
+  border(part = "all", border = officer::fp_border(color = "black", width = 1)) |> 
+  # Width and layout control
+  set_table_properties(
+    width = 1,                # Use 100% of available width
+    layout = "autofit"        # Smart column adjustment
+  ) |> 
+  # Column width specification (adjust as needed)
+  width(width = c(2.5, rep(1.5, ncol(nivolumab_data)-1))) |> 
+  # Additional professional formatting
+  align(align = "center", part = "header") |> 
+  align(align = "left", part = "body") |> 
+  padding(padding = 5, part = "all") |> 
+  fontsize(size = 10, part = "all") |> 
+  font(fontname = "Arial", part = "all")
+
+# Display the table
+nivolumab_table
+```
+
+```{r}
+#| label: tbl-pembrolizumab
+#| tbl-cap: "Pembrolizumab Reactions by Organ System"
+#| echo: FALSE
+#| warning: FALSE
+#| message: FALSE
+
+library(flextable)
+library(officer)  # Required for fp_border()
+
+# Load and prepare data
+pembro_data <- readRDS("../../results/tables/table_pembrolizumab.rds")[-c(1, 2), ]
+
+# Create and format flextable
+pembro_table <- flextable(pembro_data) |> 
+  bold(part = "header") |> 
+  border(part = "all", border = fp_border(color = "black", width = 1)) |> 
+  set_table_properties(width = 1, layout = "autofit") |> 
+  width(width = c(2.2, rep(1.5, ncol(pembro_data)-1))) |> 
+  align(align = "left", part = "body", j = 1) |> 
+  align(align = "left", part = "body", j = 2:ncol(pembro_data)) |> 
+  align(align = "center", part = "header") |> 
+  padding(padding = 5, part = "all") |> 
+  fontsize(size = 10, part = "all")
+
+# Display table
+pembro_table
+```
+
+The PRR and ROR when all irAEs are considered is greater than 1 for pembrolizumab (PRR: 1.8 (1.52, 2.14), ROR: 2.07 (1.69, 2.55)).
+Similarly, the IC is greater than 0 (0.36 (0.29, 0.44)).
+Thus, there is a safety signal associated with pembrolizumab for all irAEs.
+
+When considering individual irAEs, pembrolizumab also has a safety signal for rash (PRR: 2.11 (1.17, 3.8), ROR: 2.14 (1.18, 3.9), IC: 0.44 (0.17, 0.71).
+
+Atezolizumab has a safety signal for hyperthyroidism (PRR: 3.51 (1.07, 11.45), ROR: 3.53 (1.07, 11.61), IC:	1.1 (0.33, 1.87).
+There also may be a signal for nephritis (PRR: 3.65 (0.98, 13.56), ROR: 3.68 (0.98, 13.73), IC:	1.12 (0.29, 1.96), though the 95% CI for PRR and ROR is not greater than 1 (though very close to 1).
+
+Similarly, avelumab may has a safety signal for hypothyroidism (PRR: 3.05 (0.98, 9.51), ROR:	3.08 (0.97, 9.75), IC:	1.32 (0.12, 2.52)).
+
+No significant safety signals were found for nivolumab.
+
+A random forest model was built to distinguish between serious and non-serious outcomes using age, sex, and ICI.
+Inverse weighting was used to deal with the class imbalance.
+The model was built using cross-validation.
+
+```{r}
+#| label: tbl-randomforest
+#| tbl-cap: "Random Forest Evaluation Metrics"  # Removed "Table X" to prevent duplication
+#| echo: FALSE
+#| warning: FALSE
+#| message: FALSE
+
+library(flextable)
+library(officer)
+
+# Load and prepare data
+rf_metrics <- readRDS("../../results/tables/randomforesttable.rds")
+
+# Create professional flextable
+flextable(rf_metrics) |> 
+  # Basic formatting
+  bold(part = "header") |> 
+  border(part = "all", border = fp_border(color = "black", width = 1)) |> 
+  # Number formatting
+  colformat_double(digits = 3) |>  # Adjust decimal places as needed
+  # Alignment
+  align(align = "center", part = "header") |> 
+  align(align = "right", part = "body", j = which(sapply(rf_metrics, is.numeric))) |> 
+  # Layout
+  set_table_properties(width = 1, layout = "autofit") |> 
+  width(width = c(1.8, rep(1.2, ncol(rf_metrics)-1))) |>  # Adjust column widths
+  # Visual enhancements
+  bg(part = "header", bg = "#f2f2f2") |>  # Light gray header
+  fontsize(size = 10, part = "all") |> 
+  padding(padding = 4, part = "all")
+```
+
+We see that the accuracy is quite low (~25%).
+The model is not able to distinguish well between serious and non-serious cases, classifying most non-serious cases as serious.
+This issue arises due to the class imbalance in the data used to train the model (305 serious cases out of 365).
+Advanced machine learning methods may be explored to overcome this problem. Some additional methods will be considered in part 5.
 
@@
 
-*Summarize what you did, what you found and what it means.*
+In this study, the safety signals for the most commonly reported irAEs in clinical trials were assessed.
+Safety signals for pembrolizumab included all irAEs as a whole, as well as a signal for rash.
+Hyperthyroidism was a concern with atezolizumab, while nephritis may also be an issue.
+Hypothyroidism could be of concern with avleumab use.
+No specific signals were found with nivolumab.
 
@@
 
-*Discuss what you perceive as strengths and limitations of your analysis.*
+This study used data from the FAERS dashboard, and thus provides real-world evidence regarding the association of irAEs with ICI use in bladder cancer.
+To my knowledge, this is the first study using FAERS data to assess this association.
+However, the findings of this study must be considered in context of certain limitations.
+Firstly, FAERS data has several limitations. The dashboard cannot provide causal information. 
+This is a voluntary system. In the US, only drug manufacturers are mandated to reported adverse events. This means that all events are not reported.
+Additionally, the reports may have duplicates, or may be incomplete. Efforts have been made in this study to remove the duplicates, but data completeness cannot be ensured.
+Pembrolizumab was approved in 2014, while avelumab, nivolumab, and atezolizumab were initially approved in 2017, with additonal approvals in later years for other stages of bladder cancer.
+As such, the larger number of reports associated with pembrolizumab may simple be due to the larger reporting time.
+A future analysis with additional data may find different results to the one obtained using data until December 2024.
 
@@
 
-*What are the main take-home messages?*
+Immune checkpoint inhibitor therapy are associated with immune-related adverse events (particularly rash, hypothyroidism, hyperthyroidism, and nephritis) in bladder cancer.
+Currently, no established guidelines exist to manage these adverse events.
+There is a need to develop centralized guidelines and protocols for managing irAEs associated with ICI therapy.
 
-*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like.
-
-{{< pagebreak >}}
+*More details will be added in part 5 to the discussion section.*
 

### 2025-04-10 code/analysis-code/statistical-analysis.qmd
Updates to Part 4 of MADA Project

--- 2025-03-14 code/analysis-code/statistical-analysis.qmd
@@
 ```
+
+First, let's fit a logistic regression model to see which factors affect the severity of the reactions:
+
+Let's look at the relationship between severity and sex. We need to recode the sex variable for this:
+```{r}
+# Create a new binary column: 1 = Serious, 0 = Non-Serious
+mydata$serious_bin <- ifelse(mydata$serious == "Serious", 1, 0)
+```
+
+Now, we fit the model and look at the plot.
+
+```{r}
+library(ggplot2)
+
+logit_model_sex <- glm(serious_bin ~ sex, data = mydata, family = binomial)
+summary(logit_model_sex)
+mydata$predicted_prob <- predict(logit_model_sex, type = "response")
+
+severity_sex = ggplot(mydata, aes(x = sex, y = predicted_prob)) +
+  geom_jitter(width = 0.2, alpha = 0.3, color = "purple") +
+  labs(
+    title = "Predicted Probability of Serious Event by Sex",
+    x = "Sex",
+    y = "Predicted Probability"
+  ) +
+  theme_minimal()
+
+print(severity_sex)
+
+# Define paths
+severity_sex_path <- here("results", "figures", "severity_sex.png")
+logit_model_sex_path = here("results", "tables", "logit_model_sex.rds")
+
+# Save all four tables to their respective paths
+ggsave(filename = severity_sex_path, plot=severity_sex) 
+saveRDS(logit_model_sex, file = logit_model_sex_path)
+
+```
+
+We see that the probability of serious outcomes is highest for sex not specified, followed by males, then females.
+There are more cases for males than females, so it is expected that more males report severe cases.
+However, it is surprising that even with a low percentage of non-specified sex reports, the severity can be high.
+This may indicate underlying issues with the data reporting.
+We do see that sex is not a significant predictor of severity.
+
+Let's look at the association of age with severity.
+We know that age has a few NA's. Since, this is continuous data, it is better to remove the NA's.
+
+```{r}
+mydata_clean <- mydata[!is.na(mydata$patient_age), ]
+```
+
+```{r}
+logit_model_age <- glm(serious_bin ~ patient_age, data = mydata_clean, family = binomial)
+summary(logit_model_age)
+mydata_clean$predicted_prob <- predict(logit_model_age, type = "response")
+
+severity_age = ggplot(mydata_clean, aes(x = patient_age, y = predicted_prob)) +
+  geom_point(alpha = 0.3, color = "steelblue") +
+  labs(
+    title = "Predicted Probability of Serious Outcome by Age",
+    x = "Patient Age",
+    y = "Predicted Probability"
+  ) +
+  theme_minimal()
+
+print(severity_age)
+
+# Define paths
+severity_age_path <- here("results", "figures", "severity_age.png")
+logit_model_age_path = here("results", "tables", "logit_model_age.rds")
+
+# Save all four tables to their respective paths
+ggsave(filename = severity_age_path, plot=severity_age) 
+saveRDS(logit_model_age, file = logit_model_age_path)
+
+```
+We see that predicted severity decreases with age.
+However, this relationship is also not signficant.
+
+Now, we use machine learning methods (random forest) to assess severity of reaction based on age, sex and drug.
+We know that there are NA's in the age.
+Let's impute the data (since removing the variables will create a very small dataset).
+There were too many NA's for KNN imputation to be used.
+
+```{r}
+# Calculate the mean of the 'patient_age' column, excluding missing values
+mean_age <- mean(mydata$patient_age, na.rm = TRUE)
+
+# Replace missing values (NA) in 'patient_age' with the calculated mean
+mydata$patient_age_imputed <- ifelse(is.na(mydata$patient_age), mean_age, mydata$patient_age)
+
+```
+
+First, let's split the data:
+
+```{r}
+set.seed(123)
+library(caret)
+
+train_index <- createDataPartition(mydata$serious_bin, p = 0.7, list = FALSE)
+train_data <- mydata[train_index, ]
+test_data  <- mydata[-train_index, ]
+
+```
+
+First, let's ensure that the serious variable is a factor:
+```{r}
+# Ensure 'serious_bin' and 'sex' is a factor
+train_data$serious_bin <- as.factor(train_data$serious_bin)
+test_data$serious_bin <- as.factor(test_data$serious_bin)
+
+train_data$sex <- factor(train_data$sex)
+test_data$sex <- factor(test_data$sex)
+```
+
+First, let's use a basic logistic regression model:
+```{r}
+# Fit logistic regression model with cross-validation
+logit_model <- train(serious_bin ~ patient_age_imputed + sex + suspect_product_active_ingredients, 
+                     data = train_data, 
+                     method = "glm", 
+                     family = "binomial", 
+                     trControl = trainControl(method = "cv", number = 5))
+
+```
+
+Now, let's fit a decision tree (for classification):
+```{r}
+# Decision Tree (CART) model with cross-validation
+dt_model <- train(serious_bin ~ patient_age_imputed + sex + suspect_product_active_ingredients, 
+                  data = train_data, 
+                  method = "rpart", 
+                  trControl = trainControl(method = "cv", number = 5))
+
+```
+
+Now, let's use a random forest:
+```{r}
+# Random Forest model with cross-validation
+rf_model <- train(serious_bin ~ patient_age_imputed + sex + suspect_product_active_ingredients, 
+                  data = train_data, 
+                  method = "rf", 
+                  trControl = trainControl(method = "cv", number = 5))
+
+```
+
+Now, let's evaluate all the models:
+
+```{r}
+# Accuracy and AUC for Logistic Regression
+logit_model$results
+
+# Accuracy and AUC for Decision Tree
+dt_model$results
+
+# Accuracy and AUC for Random Forest
+rf_model$results
+
+```
+
+```{r}
+library(pROC)
+library(ggplot2)
+library(here)
+
+# Logistic Regression ROC
+logit_roc <- roc(train_data$serious_bin, predict(logit_model, type = "prob")[,2])
+logit_roc_plot <- ggroc(logit_roc) + ggtitle("Logistic Regression ROC Curve")
+cat("Logistic Regression AUC:", auc(logit_roc), "\n")
+
+# Decision Tree ROC
+dt_roc <- roc(train_data$serious_bin, predict(dt_model, type = "prob")[,2])
+dt_roc_plot <- ggroc(dt_roc) + ggtitle("Decision Tree ROC Curve")
+cat("Decision Tree AUC:", auc(dt_roc), "\n")
+
+# Random Forest ROC
+rf_roc <- roc(train_data$serious_bin, predict(rf_model, type = "prob")[,2])
+rf_roc_plot <- ggroc(rf_roc) + ggtitle("Random Forest ROC Curve")
+cat("Random Forest AUC:", auc(rf_roc), "\n")
+
+# Define file paths
+logit_roc_path <- here("results", "figures", "logit_roc.png")
+dt_roc_path <- here("results", "figures", "dt_roc.png")
+rf_roc_path <- here("results", "figures", "rf_roc.png")
+
+# Save plots
+ggsave(filename = logit_roc_path, plot = logit_roc_plot, width = 6, height = 4)
+ggsave(filename = dt_roc_path, plot = dt_roc_plot, width = 6, height = 4)
+ggsave(filename = rf_roc_path, plot = rf_roc_plot, width = 6, height = 4)
+
+```
+
+```{r}
+cat("\nModel Comparison based on AUC:\n")
+cat("Logistic Regression AUC:", auc(logit_roc), "\n")
+cat("Decision Tree AUC:", auc(dt_roc), "\n")
+cat("Random Forest AUC:", auc(rf_roc), "\n")
+```
+
+We see that the Random Forest model has an AUC of 0.806. Thus, this models is able to predict between serious and non-serious cases ~81% of the time.
+The logistic regression model is next, able to distinguish ~60% of the time.
+The decision tree model performs poorly, possibly due to fitting issues.
+
+Let's fit the random forest model to the test data:
+```{r}
+# Make predictions on the test data (probabilities for AUC)
+rf_pred_prob <- predict(rf_model, newdata = test_data, type = "prob")
+
+# Make predictions on the test data (class labels for evaluation)
+rf_pred_class <- predict(rf_model, newdata = test_data, type = "raw")
+
+# ROC curve for Random Forest
+rf_roc <- roc(test_data$serious_bin, rf_pred_prob[, 2])  # Using the probability of the 'Serious' class
+plot(rf_roc, main = "Random Forest ROC Curve")
+cat("Random Forest AUC on Test Data:", auc(rf_roc), "\n")
+
+# Extract the random forest model from the train object
+rf_final_model <- rf_model$finalModel
+
+library(randomForest)
+# Display feature importance
+importance(rf_final_model)
+
+# Plot feature importance
+varImpPlot(rf_final_model)
+
+```
+
+Save the file:
+```{r}
+# Load library
+library(randomForest)
+
+# Define file path
+importance_plot_path <- here::here("results", "figures", "rf_feature_importance.png")
+
+# Save the plot
+png(filename = importance_plot_path, width = 800, height = 600)
+varImpPlot(rf_final_model)
+dev.off()
+
+```
+
+We see that the model does not perform as well on the test data (AUC=0.56), and is able to distinguish the severity only 56% of the time.
+The most important feature is the patient age.
+However, this variable has imputations and thus may affect the dependability of the model.
+Overall, the model performs decently on train data, but not on test data.
+More data and variables can help improve the fit.
+With this data, this model is not useful in distinguishing between serious and non-serious cases.
+
+Let's create a summary table for all the models:
+```{r}
+library(pROC)
+library(caret)
+
+# Logistic Regression Model
+logit_preds <- predict(logit_model, newdata = train_data)
+logit_acc <- confusionMatrix(logit_preds, train_data$serious_bin)$overall['Accuracy']
+logit_auc <- roc(train_data$serious_bin, predict(logit_model, newdata = train_data, type = "prob")[,2])$auc
+logit_sens <- confusionMatrix(logit_preds, train_data$serious_bin)$byClass['Sensitivity']
+logit_spec <- confusionMatrix(logit_preds, train_data$serious_bin)$byClass['Specificity']
+
+# Decision Tree Model (now using dt_model)
+dt_preds <- predict(dt_model, newdata = train_data)
+dt_acc <- confusionMatrix(dt_preds, train_data$serious_bin)$overall['Accuracy']
+dt_auc <- roc(train_data$serious_bin, predict(dt_model, newdata = train_data, type = "prob")[,2])$auc
+dt_sens <- confusionMatrix(dt_preds, train_data$serious_bin)$byClass['Sensitivity']
+dt_spec <- confusionMatrix(dt_preds, train_data$serious_bin)$byClass['Specificity']
+
+# Random Forest Model
+rf_preds <- predict(rf_model, newdata = train_data)
+rf_acc <- confusionMatrix(rf_preds, train_data$serious_bin)$overall['Accuracy']
+rf_auc <- roc(train_data$serious_bin, predict(rf_model, newdata = train_data, type = "prob")[,2])$auc
+rf_sens <- confusionMatrix(rf_preds, train_data$serious_bin)$byClass['Sensitivity']
+rf_spec <- confusionMatrix(rf_preds, train_data$serious_bin)$byClass['Specificity']
+
+# Create summary data frame with results
+model_summary <- data.frame(
+  Model = c("Logistic Regression", "Decision Tree", "Random Forest"),
+  Accuracy = c(logit_acc, dt_acc, rf_acc),
+  AUC = c(logit_auc, dt_auc, rf_auc),
+  Sensitivity = c(logit_sens, dt_sens, rf_sens),
+  Specificity = c(logit_spec, dt_spec, rf_spec)
+)
+
+# Convert to huxtable
+model_table <- as_hux(model_summary)
+
+# Format the table
+model_table <- set_bold(model_table, 1, 1:ncol(model_table))  # Make the entire header row bold
+model_table <- set_align(model_table, 1, 1:ncol(model_table), "center")  # Center-align the entire header row
+model_table <- set_all_borders(model_table, 1)
+
+# Print the table
+model_table
+
+# Save the table
+model_table_path = here("results", "tables", "model_table.rds")
+saveRDS(model_table, file = model_table_path)
+```
+We see that sensitivity is 0 and specificity is high, meaning all the models always predict 'serious'.
+Such problems arise due to class imbalances, and may require advanced methods to deal with.

### 2025-04-10 code/eda-code/eda.qmd
Updates to Part 4 of MADA Project

@@
 
-# Convert to huxtable object
-huxtable_table <- as_hux(transposed_table)
+# Get the current order of attributes
+current_order <- transposed_table$Attribute
 
-# Format the table with huxtable (style, borders, etc.)
+# Create a new order with subheadings inserted
+new_order <- c()
+insert_subheading <- function(order_vector, before, subheading) {
+  pos <- which(order_vector == before)
+  if (length(pos) > 0) {
+    c(order_vector[1:(pos-1)], subheading, order_vector[pos:length(order_vector)])
+  } else {
+    order_vector
+  }
+}
+
+# Insert subheadings before specific attributes
+new_order <- insert_subheading(current_order, "Male", "Sex")
+new_order <- insert_subheading(new_order, "Serious", "Severity of reaction")
+new_order <- insert_subheading(new_order, "Died", "Outcomes")
+new_order <- insert_subheading(new_order, "Healthcare Professional", "Reporter type")
+new_order <- insert_subheading(new_order, "US", "Country")
+
+# Create a new table with subheadings
+subheading_rows <- data.frame(
+  Attribute = setdiff(new_order, current_order),
+  stringsAsFactors = FALSE
+)
+subheading_rows[colnames(transposed_table)[-1]] <- ""
+
+final_table <- bind_rows(
+  lapply(new_order, function(x) {
+    if (x %in% subheading_rows$Attribute) {
+      subheading_rows[subheading_rows$Attribute == x, ]
+    } else {
+      transposed_table[transposed_table$Attribute == x, ]
+    }
+  })
+)
+
+# Create the final table with title and clean formatting
+huxtable_table <- as_hux(final_table, add_colnames = TRUE) %>% 
+  # Add a title row
+  insert_row(c("Adverse Event Report Summary", rep("", ncol(final_table)-1)), 
+             after = 0) %>% 
+  # Merge title cells
+  merge_cells(1, 1:ncol(final_table)) %>%
+  # Basic formatting
+  set_all_borders(1) %>%               # All cell borders
+  set_align(everywhere, everywhere, "center") %>%  # Center alignment
+  set_valign(everywhere, everywhere, "middle") %>% # Vertical center
+  set_font_size(10) %>%               # Base font size
+  set_font_size(1, everywhere, 12) %>% # Larger font for title
+  set_bold(1, everywhere, TRUE) %>%   # Bold title only
+  # Ensure no other text is bolded
+  set_bold(2:nrow(final_table)+1, everywhere, FALSE)
+
+# Format the title row
 huxtable_table <- huxtable_table %>%
-  set_all_borders(1) %>%
-  set_font_size(12) %>%
-  set_align("center") %>%
-  set_width(1.5)
+  set_background_color(1, everywhere, "gray95") %>%
+  set_top_border(1, everywhere, 2) %>%
+  set_bottom_border(1, everywhere, 2)
 
-# Print the huxtable object, this will render correctly in Word
 huxtable_table
@@
 
-# Add styling to the table
-ht <- ht %>%
-    set_bold(1, 1:ncol(ht)) %>%  # Bold the first row (header)
-    set_caption("Immune Checkpoint Inhibitors by Organ System and Reaction") %>%
-    set_width(1)                  # Set column width to make the table more compact
+# FINAL WORKING SOLUTION
+ht <- as_hux(organ_system_pivoted, add_colnames = TRUE) %>%
+  # Basic formatting
+  set_bold(1, everywhere, TRUE) %>%      # Bold headers
+  set_align(everywhere, everywhere, "center") %>%
+  set_all_borders(brdr(1, "solid")) %>%  # Solid borders
+  set_all_padding(4) %>%
+  set_font_size(10) %>%
+  set_width(0.9)
 
-# Display the table
+# Add title as a caption (works in exports)
+caption(ht) <- "Immune Checkpoint Inhibitors by Organ System and Reaction"
+
+# Bold specific drug names
+drug_cols <- which(colnames(organ_system_pivoted) %in% 
+                  c("atezolizumab", "pembrolizumab", "nivolumab", "avelumab"))
+if (length(drug_cols) > 0) {
+  ht <- ht %>% set_bold(1, drug_cols, TRUE)
+}
+
+# Display
 ht
@@
 ```
-
 # Data exploration through figures

### 2025-04-10 products/manuscript/Manuscript.qmd
Updates to Part 4 of MADA Project

@@
 
+{{< pagebreak >}}
+
 ## Description of data and data source
@@
 To assess whether an ICI has a safety signal associated with a particular irAE, three metrics were used:
+
 1. Proportional reporting ratio (PRR):
@@
 A random forest model was fit to distinguish between serious and non-serious cases on the basis of age, sex, and ICI.
+
+{{< pagebreak >}}
 
@@
 ```
+{{< pagebreak >}}
 
@@
 This was followed by creating some descriptive figures.
+Two logistic regression models were fit to assess the association by severity with sex and age.
 The PRR, ROR and IC (along with their 95% CI) were then calculated to assess the drug signal.
-A random forest model was built using age, sex, and ICI, to distinguish between serious and non-serious cases.
+Three different machine learning models were used to distinguish severity on the bases of sex, age, and type of ICI.
+Due to missing values, the patient age values were imputed.
+First, the data was split into training and testing data (70/30 split). A 5-fold cross-validation was used.
 
@@
 
-## Basic statistical analysis
+{{< pagebreak >}}
 
@@
 ```
+
 Pembrolizumab has the highest number of reactions.
@@
 These include reactions which do not belong to the other organ classes, such as myositis.
+
+{{< pagebreak >}}
+
+## Basic statistical analysis
+
+@fig-result shows the number of reactions by drug, along with the seriousness of the reaction.
+
+```{r}
+#| label: fig-result
+#| fig-cap: "Association of severity with sex"
+#| echo: FALSE
+library(here)
+library(knitr)
+knitr::include_graphics(here("results","figures","severity_sex.png"))
+```
+
+The probability of serious outcomes is highest for sex not specified, followed by males, then females.
+There are more cases for males than females, so it is expected that more males report severe cases.
+However, it is surprising that even with a low percentage of non-specified sex reports, the severity can be high.
+This may indicate underlying issues with the data reporting.
+However, sex is not a significant predictor of severity.
+
+{{< pagebreak >}}
+
+@fig-result2 shows the association between severity and age.
+
+```{r}
+#| label: fig-result2
+#| fig-cap: "Association of severity with age"
+#| echo: FALSE
+library(here)
+library(knitr)
+knitr::include_graphics(here("results","figures","severity_age.png"))
+```
+
+The predicted severity decreases with age.
+However, this relationship is also not signficant.
+
+{{< pagebreak >}}
 
@@
 
+{{< pagebreak >}}
+
 ```{r}
@@
 ```
+
+{{< pagebreak >}}
 
@@
 
-A random forest model was built to distinguish between serious and non-serious outcomes using age, sex, and ICI.
-Inverse weighting was used to deal with the class imbalance.
-The model was built using cross-validation.
+{{< pagebreak >}}
+
+## Machine Learning Analysis
+
+@tbl--randomforest shows the results of the three machine learning models (logistic regression, decision tree, random forest).
 
@@
 # Load and prepare data
-rf_metrics <- readRDS("../../results/tables/randomforesttable.rds")
+rf_metrics <- readRDS("../../results/tables/model_table.rds")
 
-# Create professional flextable
-flextable(rf_metrics) |> 
-  # Basic formatting
+rf_metrics_cleaned <- rf_metrics[-1, ]  # Removes the second row
+
+flextable(rf_metrics_cleaned) |> 
   bold(part = "header") |> 
   border(part = "all", border = fp_border(color = "black", width = 1)) |> 
-  # Number formatting
-  colformat_double(digits = 3) |>  # Adjust decimal places as needed
-  # Alignment
+  colformat_double(digits = 3) |>  
   align(align = "center", part = "header") |> 
-  align(align = "right", part = "body", j = which(sapply(rf_metrics, is.numeric))) |> 
-  # Layout
+  align(align = "right", part = "body", j = which(sapply(rf_metrics_cleaned, is.numeric))) |> 
   set_table_properties(width = 1, layout = "autofit") |> 
-  width(width = c(1.8, rep(1.2, ncol(rf_metrics)-1))) |>  # Adjust column widths
-  # Visual enhancements
-  bg(part = "header", bg = "#f2f2f2") |>  # Light gray header
+  width(width = c(1.8, rep(1.2, ncol(rf_metrics_cleaned)-1))) |>  
+  bg(part = "header", bg = "#f2f2f2") |>  
   fontsize(size = 10, part = "all") |> 
   padding(padding = 4, part = "all")
+
 ```
 
-We see that the accuracy is quite low (~25%).
-The model is not able to distinguish well between serious and non-serious cases, classifying most non-serious cases as serious.
-This issue arises due to the class imbalance in the data used to train the model (305 serious cases out of 365).
-Advanced machine learning methods may be explored to overcome this problem. Some additional methods will be considered in part 5.
+The Random Forest model has an AUC of 0.806. Thus, this models is able to predict between serious and non-serious cases ~81% of the time.
+The logistic regression model is next, able to distinguish ~60% of the time.
+The decision tree model performs poorly, possibly due to fitting issues.
+All the models have the same accuracy. Additionally, sensitivity is 0 and specificity is high, meaning all the models always predict 'serious'.
+Such problems arise due to class imbalances, and may require advanced methods to deal with.
+The random forest model does not perform as well on the test data (AUC=0.56), and is able to distinguish the severity only 56% of the time.
+
+@fig-result3 shows the feature importance.
+
+```{r}
+#| label: fig-result3
+#| fig-cap: "Feature importance"
+#| echo: FALSE
+library(here)
+library(knitr)
+knitr::include_graphics(here("results","figures","rf_feature_importance.png"))
+```
+
+The most important feature is the patient age.
+However, this variable has imputations and thus may affect the dependability of the model.
+Overall, the model performs decently on train data, but not on test data.
+More data and variables can help improve the fit.
+With this data, this model is not useful in distinguishing between serious and non-serious cases.
+The ROC curves for the different models are available in the Supplementary Material.
 

### 2025-04-10 products/manuscript/supplement/Supplementary-Material.qmd
Updates to Part 4 of MADA Project

---
title: "Supplement to Example Manuscript Template for a Data Analysis Project"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../../assets/dataanalysis-references.bib
csl: ../../../assets/nlm-style.csl
---

# Overview

This supplement has a few exploratory figures and discussion of additional methods that can be used for disproportionality analysis.

# Code and file information

To reproduce this study:
First run the processing code file. This processed the raw data in the 'data' subfolder and saves it as an rds file.
Then, run the exploratory data (eda) code file. This file produces summary tables and figures.
The figures are available in this supplement.
Then, run the analysis file. This conducts a disproportionality analysis and runs a random forest model to assess serious vs non-serious cases based on age, sex, and ICI.

{{< pagebreak >}}

# Additional Method Details

This part will be updated in part 5 - after improving the random forest code.
{{< pagebreak >}}

# Additional results

@fig-result shows the number of reactions by drug, along with the seriousness of the reaction.

```{r}
#| label: fig-result
#| fig-cap: "Bar chart of outcome of immune checkpoint inhibitors"
#| echo: FALSE
library(here)
library(knitr)
knitr::include_graphics(here("results","figures","figure1.png"))
```

@fig-result1 shows the bar chart by type of organ system involved in the reaction.

```{r}
#| label: fig-result1
#| fig-cap: "Bar chart of types of reactions"
#| echo: FALSE
knitr::include_graphics(here("results","figures","figure2.png"))
```

@fig-result2 shows the bar chart by type of organ system involved in the reaction.

```{r}
#| label: fig-result2
#| fig-cap: "ROC Curve - Logistic Regression"
#| echo: FALSE
knitr::include_graphics(here("results","figures","logit_roc.png"))
```

@fig-result3 shows the bar chart by type of organ system involved in the reaction.

```{r}
#| label: fig-result3
#| fig-cap: "ROC Curve - Decision Tree"
#| echo: FALSE
knitr::include_graphics(here("results","figures","dt_roc.png"))
```

@fig-result4 shows the bar chart by type of organ system involved in the reaction.

```{r}
#| label: fig-result4
#| fig-cap: "ROC Curve - Random forest"
#| echo: FALSE
knitr::include_graphics(here("results","figures","rf_roc.png"))
```

# Discussion

Additional discussion will be done in part 5 after improving the random effects model.
[@Leslie2025-lp] is a placeholder reference.

{{< pagebreak >}}

# References

### 2025-04-18 products/manuscript/Manuscript.qmd
Part 5 of MADA Project

@@
 ---
-title: "Immune-Related Adverse Events (irAEs) associated with Immune Checkpoint Inhibitor (ICI) Therapy in Bladder Cancer: An Analysis of the FDA Adverse Event Reporting System (FAERS)"
+title: "Immune-Related Adverse Events (irAEs) Associated with Immune Checkpoint Inhibitor (ICI) Therapy in Bladder Cancer: An Analysis of the FDA Adverse Event Reporting System (FAERS)"
 author: "Pooja Gokhale, PharmD"
@@
 Background: Bladder cancer is the ninth most common type of cancer worldwide.
-Treatment options include surgery, chemotherapy and immunotherapy.
+Treatment options include surgery, chemotherapy, and immunotherapy.
 Immune checkpoint inhibitors (ICIs) are a class of immunotherapeutic agents recommended for the treatment of bladder cancer.
@@
 
-Methods: Reports in the FAERS dashbaord through December 2024 were analyzed for irAEs to four ICIs recommended in bladder cancer - atezolizumab, avelumab, nivolumab, and pembrolizumab.
+Methods: Reports in the FAERS dashboard through December 2024 were analyzed for irAEs to four ICIs recommended in bladder cancer: atezolizumab, avelumab, nivolumab, and pembrolizumab.
 Standardized queries were used to identify cases of rash, pruritus, hypothyroidism, hyperthyroidism, colitis, hepatitis, and nephritis.
@@
 Results: Out of 2948 cases identified from FAERS, 365 cases with irAEs were analyzed.
-The largest number of cases were reported for pembrolizumab (n=307), followed by atezolizumab (n = 91), avelumab (n = 41), and nivolumab (n =37).
+The largest number of cases was reported for pembrolizumab (n=307), followed by atezolizumab (n = 91), avelumab (n = 41), and nivolumab (n =37).
 Pembrolizumab was found to have safety signals for all irAEs considered together (PRR: 1.8 (1.52, 2.14), ROR: 2.07 (1.69, 2.55), IC:	0.36 (0.29, 0.44)).
@@
 Bladder cancer is the ninth most common type of cancer, according to a 2022 report published by the International Agency for Research on Cancer [@Bray2024].
-The American Cancer Society estimates approximately 85,000 new bladder cancer cases, and approximately 17,000 bladder cancer deaths in the United States in 2025.
+The American Cancer Society estimates approximately 85,000 new bladder cancer cases and approximately 17,000 bladder cancer deaths in the United States in 2025.
 
@@
 According to the National Comprehensive Cancer Network (NCCN) guidelines, immunotherapy with immune checkpoint inhibitors (ICIs) is recommended at various stages of bladder cancer treatment.
-ICIs such as pembrolizumab, atezolizumab, nivolumab, and avelumab are FDA approved and recommended for the adjuvant and first and second-line treatment of bladder cancer [@Flaig2024].
+ICIs such as pembrolizumab, atezolizumab, nivolumab, and avelumab are FDA-approved and recommended for the adjuvant and first and second-line treatment of bladder cancer [@Flaig2024].
 
-ICIs are associated with various adverse effects, particularly with immune-related adverse events (irAEs) such as pneumonitis, hepatitis, colitis, hypothyroidism, dermatitis, and others [@Yin2023].This study aims to analyze real-world adverse event data from the FDA Adverse Event Reporting System associated with ICI therapy in bladder cancer.
+ICIs are associated with various adverse effects, particularly with immune-related adverse events (irAEs) such as pneumonitis, hepatitis, colitis, hypothyroidism, dermatitis, and others [@Yin2023].
+This study aims to analyze real-world adverse event data from the FDA Adverse Event Reporting System associated with ICI therapy in bladder cancer.
 
@@
 The study aims to analyze whether irAEs associated with ICIs generate a signal of disproportionate reporting (SDR).
-The goal is to determine whether irAEs associated with ICIs in baldder cancer are greater than expected, and thus further research and strategies to deal with these adverse events is required.
+The goal is to determine whether irAEs associated with ICIs in bladder cancer are greater than expected, and thus, further research and strategies to deal with these adverse events are required.
 
@@
 
-A random forest model was fit to distinguish between serious and non-serious cases on the basis of age, sex, and ICI.
+The association between severity of reactions and age and sex was assessed using logistic regression.
+Three machine learning models (logistic regression, decision tree, and random forest) were used to distinguish between serious and non-serious reactions.
 
@@
 
-## Data aquisition
+## Data acquisition
 
@@
 Pembrolizumab, atezolizumab, avelumab, and nivolumab were the keywords used.
-Previously, durvalumab, and ipilimumab were also approved for bladder cancer treatment.
+Previously, durvalumab and ipilimumab were also approved for bladder cancer treatment.
 However, this indication has since then been removed by the manufacturers.
@@
 The file was downloaded as an Excel file.
-First, the column were renamed so as to make them easier to use in code.
-Then, products with concomitant drugs were removed, so as to focus only on cases associated with ICIs.
-Then, cases which also had other drugs as the suspect active ingredients were removed.
+Then, in R, the columns were renamed so as to make them easier to use in code.
+Then, products with concomitant drugs were removed so as to focus only on cases associated with ICIs.
+Then, cases that also had other drugs as the suspect active ingredients were removed.
 This was followed by retaining only the irAEs.
@@
 This list is available in the supplementary material.
-Any reactions which had the same age, weight, FDA received date, country of occurrence, suspected active ingredient, and reported reaction were considered duplicates and were removed.
-Following this, the reactions were coded according to the MedDRA Preferred Terms (PTs), and classified according to the System Organ Class (SOC).
-The age values were converted to numeric and the countries were re-classified as US, outside US, and not specified.
+This list was informed by common irAEs reported in various clinical trials  [@Galsky2020] [@Galsky2020] [@Bajorin2021] [@Powles2020].
+Any reactions that had the same age, weight, FDA received date, country of occurrence, suspected active ingredient, and reported reaction were considered duplicates and were removed.
+Following this, the reactions were coded according to the MedDRA Preferred Terms (PTs) and classified according to the System Organ Class (SOC).
+The age values were converted to numeric, and the countries were reclassified as US, outside US, and not specified.
 The code for the analysis would involve loading the file, removing duplicates, assessing missing data, and deciding whether to remove it. The details are available in the processing code file.
@@
 Two logistic regression models were fit to assess the association by severity with sex and age.
-The PRR, ROR and IC (along with their 95% CI) were then calculated to assess the drug signal.
-Three different machine learning models were used to distinguish severity on the bases of sex, age, and type of ICI.
+The PRR, ROR, and IC (along with their 95% CI) were then calculated to assess the drug signal.
+Three different machine learning models were used to distinguish severity on the basis of sex, age, and type of ICI.
 Due to missing values, the patient age values were imputed.
@@
 
-Most of the reaction were reported by healthcare professionals. Since FAERS is a US-based dashboard, many reactions were reported from the US.
+Most of the reactions were reported by healthcare professionals. Since FAERS is a US-based dashboard, many reactions were reported from the US.
 
@@
 The largest percentage of reactions are classified as "Other".
-These include reactions which do not belong to the other organ classes, such as myositis.
+These include reactions that do not belong to the other organ classes, such as myositis.
 
@@
 The predicted severity decreases with age.
-However, this relationship is also not signficant.
+However, this relationship is also not significant.
 
@@
 
-The following tables show the results from the disproportionality analysis.
+The following tables show the results of the disproportionality analysis.
 
@@
 
-Similarly, avelumab may has a safety signal for hypothyroidism (PRR: 3.05 (0.98, 9.51), ROR:	3.08 (0.97, 9.75), IC:	1.32 (0.12, 2.52)).
+Similarly, avelumab may have a safety signal for hypothyroidism (PRR: 3.05 (0.98, 9.51), ROR:	3.08 (0.97, 9.75), IC:	1.32 (0.12, 2.52)).
 
@@
 
-@tbl--randomforest shows the results of the three machine learning models (logistic regression, decision tree, random forest).
+@tbl-randomforest shows the results of the three machine learning models (logistic regression, decision tree, random forest).
 
@@
 
+library(dplyr)
 library(flextable)
-library(officer)
 
-# Load and prepare data
+# Load and clean data
 rf_metrics <- readRDS("../../results/tables/model_table.rds")
+rf_metrics_cleaned <- rf_metrics[-1, ]
 
-rf_metrics_cleaned <- rf_metrics[-1, ]  # Removes the second row
+# Convert numeric-looking columns to double (if needed)
+rf_metrics_cleaned <- rf_metrics_cleaned %>%
+  mutate(across(where(~ all(grepl("^[-+]?[0-9]*\\.?[0-9]+$", .))), as.numeric))
 
+# Build styled flextable
 flextable(rf_metrics_cleaned) |> 
@@
   border(part = "all", border = fp_border(color = "black", width = 1)) |> 
-  colformat_double(digits = 3) |>  
+  colformat_double(digits = 3, na_str = "-") |>  
   align(align = "center", part = "header") |> 
-  align(align = "right", part = "body", j = which(sapply(rf_metrics_cleaned, is.numeric))) |> 
+  align(
+    align = "left",
+    part = "body",
+    j = which(sapply(rf_metrics_cleaned, is.numeric))
+  ) |> 
   set_table_properties(width = 1, layout = "autofit") |> 
-  width(width = c(1.8, rep(1.2, ncol(rf_metrics_cleaned)-1))) |>  
+  width(width = c(1.8, rep(1.2, ncol(rf_metrics_cleaned) - 1))) |>  
   bg(part = "header", bg = "#f2f2f2") |>  
@@
 
-The Random Forest model has an AUC of 0.806. Thus, this models is able to predict between serious and non-serious cases ~81% of the time.
+The Random Forest model has an AUC of 0.806. Thus, this model is able to predict between serious and non-serious cases ~81% of the time.
 The logistic regression model is next, able to distinguish ~60% of the time.
 The decision tree model performs poorly, possibly due to fitting issues.
-All the models have the same accuracy. Additionally, sensitivity is 0 and specificity is high, meaning all the models always predict 'serious'.
+All the models have the same accuracy. Additionally, sensitivity is 0, and specificity is high, meaning all the models always predict 'serious'.
 Such problems arise due to class imbalances, and may require advanced methods to deal with.
@@
 
-*More details will be added in part 5 to the discussion section.*
+{{< pagebreak >}}
 

### 2025-04-18 products/manuscript/supplement/Supplementary-Material.qmd
Part 5 of MADA Project

@@
 
-This part will be updated in part 5 - after improving the random forest code.
+Three machine learning models were fit - logistic regression, decision tree, and random forest.
+We found that the models were not able to distinguish adequately between the serious and non-serious reactions.
+This most probably occurs due to the class imbalance - 305 serious, 60 non-serious cases.
+
 {{< pagebreak >}}
@@
 ```
+We can see that a large proportion of the reactions are serious.
 
@@
 
-@fig-result2 shows the bar chart by type of organ system involved in the reaction.
+The most common reactions are dermatological, followed by gastrointestinal.
+
+@fig-result2 shows the ROC curve for the logistic regression model.
 
@@
 ```
+We see that the curve covers about 60% of the area.
 
-@fig-result3 shows the bar chart by type of organ system involved in the reaction.
+@fig-result3 shows the ROC curve for the decision tree model.
 
@@
 
-@fig-result4 shows the bar chart by type of organ system involved in the reaction.
+The curve is a staright diagonal line (covering about 50% of the area). Thus, this model is not able to distinguish between serious and non-serious cases better than at random.
+
+@fig-result4 shows the ROC curve for the random forest model.
 
@@
 
+We see that the curve covers about 80% of the area.
+
 # Discussion
 
-Additional discussion will be done in part 5 after improving the random effects model.
-[@Leslie2025-lp] is a placeholder reference.
-
-{{< pagebreak >}}
-
-# References
+Overall, the random forest model performs better.
+However, all three models are not able to distinguish between serious and non-serious reactions, possibly due to the class imbalance.

### 2025-04-23 products/manuscript/supplement/Supplementary-Material.qmd
Part 5 of MADA Project

@@
 {{< pagebreak >}}
+
+# List of irAEs:
+Skin and subcutaneous tissue disorders:
+Rash, pruritus, dermatitis, psoriasis
+Gastrointestinal disorders:
+Colitis, hepatitis, pancreatitis, enterocolitis, ileitis, ileocolitis
+Nervous system disorders:
+Myasthenia gravis, Graves disease, myelitis, encephalitis, neuritis, peripheral neuropathy
+Endocrine disorders:
+Hyperthyroidism, hypothyroidism, thyroiditis, thyrotoxicosis, adrenitis, adrenal insufficiency, hypogonadism, hyperparathyroidism, hypercortisolism, hypophysitis, hypoparathyroidism, hypocortisolism, autoimmune diabetes
+Cardiac disorders: myocarditis, pericarditis, myocarditis
+Respiratory, thoracic, and mediastinal_disorders:
+Pneumonitis
+Renal and urinary_disorders: 
+Pyelonephritis, nephritis
 

### 2025-04-24 reviews/project-review-bydoreen.qmd
Added review 1

---
title: Project Review Template 
author: "Pooja Gokhale"
date: 04/23/2025
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Immune-Related Adverse Events (irAEs) Associated with Immune Checkpoint Inhibitor (ICI) Therapy in Bladder Cancer: An Analysis of the FDA Adverse Event Reporting System (FAERS)

Name of project author(s):

Name of project reviewer:Doreen_Kibuule_Kalembe

# Instructions

Write your comments and feedback below for each section/component of the project. The goal should be to help the author improve their project. Make comments as constructive and actionable as possible. You can provide both criticism and praise.

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

Make sure your final document compiles/renders into a readable, well-formatted html document.

Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The author described the project well. The background was well written with a series of previous work and it was cleary highlighted why they decided to take on this question.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The question was  is clearly defined in the background clearly stating the project.

### Summary assessment

* question/hypotheses was well explained

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data source was clearly stated and the all concerns regarding the data set was well explained.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The data exploratory analysis was well done and and meaningful results were provided by the author

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The analyisis was clearly done and well explained for the data that was given. All parts of data wrangling and cleaning were clearly done and each step was well explained.

### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The results were well presented with figures and tables easy to read and understand. All the figures and tables were publication quality.

### Summary assessment

* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments
The limitation of the study were clearly stated by the author and the strength were well highlighted. However the findings were not clearly discussed. Either it was short since i had a hard time understanding what your results mean. Maybe i needed to read the manuscript again but it would be better if you improved alittle bit on the discussion

### Summary assessment

*Well formulated and conclusion too.

## Further comments

The project was well scripted and well referenced. You have done well in the data exploration and analysis. Well done Pooja

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project is mostly well structured.

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The project was well documented and providing everything needed for reproduciblility.
Enough comment were given to each line of code and i was able to follow through with what the author was doing.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

All the codes are clearly provided  for reproduciblity. Project was well documented and i was able to rerun the whole analysis without having to do any edits.

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The overall study was interesting to follow through. Not many alternatives were discussed but guidelines for future research were given

### Summary assessment

* strong level of thorougness

## Further comments

Well done Pooja. Apart from the needing to add in more information about your discussion, I think your project was well scripted and I wish you the very best.

### 2025-04-28 project-review-nasir.qmd

---
title: Project Review Template 
author: YOUR NAME
date: date-modified
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Immune-Related Adverse Events (irAEs) Associated with Immune Checkpoint Inhibitor (ICI) Therapy in Bladder Cancer: An Analysis of the FDA Adverse Event Reporting System (FAERS)

Name of project author(s):Pooja Gokhale

Name of project reviewer:Muhammad Nasir

## Background, Context and Motivation

### Feedback and Comments

The study provides very clear background, strong motivation and context. The backgroud is sopported by various sources.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description

### Feedback and Comments

The study clearly state aims of the study and hypotheses

### Summary assessment

* question/hypotheses fully clear

## Data description

### Feedback and Comments

The data source is provided. The author also explain how to obtain the data, keywords used to search and filter the dataset from website. A flow process diagram also is provided, which is very helpful and easy to understand by readers.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis

### Feedback and Comments
The author explained the data cleaning and data processing in very clear way and reasonable. However, it might be useful to put more note in the code, so other people can understand and follow the process easily

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

### Feedback and Comments

The study used three different metric to assess the safety of a particular irAE: Proportional Reporting Ratio (PRR), Reporting Odd Ratios (ROR), Information Component (IC). Those are very rationable to use in this study to answer the research questions.

### Summary assessment

* strong and reasonable analysis

## Presentation

### Feedback and Comments
The author presents the result very clear and simple. It is easy to understand. All figures are interpreted and described very well.

### Summary assessment

* results are very well presented

## Discussion/Conclusions

### Feedback and Comments

There is a need to add more discussion, compare the result with other studies and use citation. The study provided enough information about strenght and limitation of the study.

### Summary assessment

* strong, complete and clear discussion

## Further comments

# Overall project content evaluation

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project is well structured. Folders and files are well placed and can connect with the manuscript. Some readme.dm files need to be updated.

### Summary assessment

* well structured

## Documentation

### Feedback and Comments

Documentation is very good. However, it would be better to add more note in the code to help reader understand in every steps.

### Summary assessment

* fully and well documented

## Reproducibility

### Feedback and Comments

All codes work well. No error occured. The file can be rendered. Well done. It is fully reproductibel.

### Summary assessment

* fully reproducible without issues

## Thoroughness

### Feedback and Comments

This is very thorough study. The author paid attention in every important step.

### Summary assessment

* strong level of thorougness

## Further comments

### 2025-05-01 products/manuscript/Manuscript.qmd
Part 6 of MADA project

@@
 
-$\land$ Corresponding author: poojaGokhale\@uga.edu
+$^{*}$ Corresponding author: poojaGokhale\@uga.edu
 
@@
 This list is available in the supplementary material.
-This list was informed by common irAEs reported in various clinical trials  [@Galsky2020] [@Galsky2020] [@Bajorin2021] [@Powles2020].
+This list was informed by common irAEs reported in various clinical trials  [@Galsky2020a] [@Galsky2020b] [@Bajorin2021] [@Powles2020].
 Any reactions that had the same age, weight, FDA received date, country of occurrence, suspected active ingredient, and reported reaction were considered duplicates and were removed.
@@
 In this study, the safety signals for the most commonly reported irAEs in clinical trials were assessed.
-Safety signals for pembrolizumab included all irAEs as a whole, as well as a signal for rash.
-Hyperthyroidism was a concern with atezolizumab, while nephritis may also be an issue.
-Hypothyroidism could be of concern with avleumab use.
-No specific signals were found with nivolumab.
+
+Pembrolizumab had a safety signal for all irAEs considered together, as well as for rash.
+
+Five clinical trials [@Galsky2020a,@Powles2021, @Apolo2025, @Bellmunt2017, @Zhang2020] have reported rash as an adverse effect associated with pembrolizumab monotherapy in bladder cancer.
+The report of rash from both clinical trials and real-world data suggests that this is a significant adverse event. 
+Rash associated with ICIs may be of various grades.
+For a grade 1 rash, the ICI therapy may be continued.
+Grade 2 rash may necessitate discontinuation of ICIs, and grade 3 rash requires discontinuation.
+Based on the grade, treatment with topical or systemic corticosteroids is appropriate.
+Dermatological consultation is usually suggested  [@Si2019].
+
+Atezolizumab had a safety signal for hyperthyroidism. Two clinical trials  [@Galsky2020b, @Bellmunt2021] reported hyperthyroidism associated with atezolizumab monotherapy in bladder cancer.
+
+Avelumab may have a safety signal for hypothyroidism. One clinical trial assessing avelumab monotherapy has reported hypothyroidism as an adverse event  [@Powles2020]. 
+Thyroid disorders are seen with about 1-15% of patients receiving ICIs  [@Ferrari2018].
+Risk factors for thyroid dysfunction with ICIs include gender (females are at a higher risk), age (younger age), race (Caucasian and Hispanic individuals are at higher risk), obesity, higher number of treatment cycles, and higher baseline thyroid-stimulating hormone (TSH)  [@Chera2022].
+
+It has been suggested that a combination of genetic factors, cellular autoimmunity, and humoral immunity may lead to thyroid dysfunction  [@Deligiorgi2021].
+ICI therapy should involve thyroid function tests every 4-6 weeks. Treatment is largely symptomatic and may involve levothyroxine therapy in the case of hypothyroidism  [@Yamauchi2025].
+
+Atezolizumab may also have a safety signal for nephritis. Two clinical trials have reported nephritis associated with atezolizumab monotherapy in bladder cancer  [@Galsky2020b, @Bellmunt2021].
+Nephritis is not a common adverse event associated with ICIs, affecting about 2-5% of patients.
+The mechanism for the development of this adverse event is not well established.
+Most patients recover with appropriate management.
+Treatment may involve discontinuation of the ICI, and corticosteroid therapy  [@Barbir2024].
+
+No prospective trials have defined strategies for effectively managing specific immune-related adverse events; thus, clinical practice remains variable.
+Committees are working to improve guidelines and to harmonize expert consensus  [@Postow2018].
+In conclusion, irAEs are associated with ICI monotherapy with bladder cancer.
+More comprehensive guidelines are required for the management of these adverse events.
+Clinicians should be aware of the potential for such adverse events.
 
